---
title: "01_01_Allometry_check"
author: "Vera Weisbecker"
date: "2024-05-03"
output: html_document
editor_options:
  chunk_output_type: console
---

##We are mostly ignoring the effect of allometry on shape in the main ms because it is not expected to have a substantial effect given the diverse sample. Here are all the analyses supporting this assumption.

```{r}

library(geomorph)  # CRAN v4.0.10 
library(ape)       # CRAN v5.8-1 
library(caper)     # CRAN v1.0.4 
library(tidyverse) # CRAN v2.0.0 
library(svglite)   # CRAN v2.2.2 
library(here)      # CRAN v1.0.2 
library(rgl)       #CRAN v1.3.34

load(here("Data", "Processed", "skullLmkSymm.rda"))
load(here("Data", "Processed", "rostralLmkSymm.rda"))
load(here("Data", "Processed", "neuroLmkSymm.rda"))
load(here("Data", "Processed", "fnpLmkSymm.rda"))
load(here("Data", "Processed", "mesLmkSymm.rda"))
load(here("Data", "Processed", "paLmkSymm.rda"))
load(file = here("Data", "Processed", "skullGPA.rda"))
load(file = here("Data", "Processed", "rostralGPA.rda"))
load(file = here("Data", "Processed", "neuroGPA.rda"))
load(file = here("Data", "Processed", "fnpGPA.rda"))
load(file = here("Data", "Processed", "mesGPA.rda"))
load(file = here("Data", "Processed", "paGPA.rda"))


# Load the labels produced in "GMM_GPA.Rmd"
pointLabels <- readRDS(here("Data", "Processed", "pointLabels.rds"))

#Load phylogeny - this is the composite tree from Rovinsky et al., 2021, with the fossil taxa added
# NB: Prothylacinus patagonicus is not included this tree
rawTree <- read.tree(here::here("Data", "phyloGMM.tre"))
```



#Allometry with Andrewsarchus is strong but related to Andrewsarchus being a high-leverage outlier.
```{r}
#shape

new.subset <-coords.subset(A = skullLmkSymm$symm.shape,
                           group = pointLabels$Taxon) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL

#reorder coords to fit the tree
meanCoords <- meanCoords[,,order(match(dimnames(meanCoords)[[3]],rawTree$tip.label))]


Csize <- as.numeric(skullGPA$Csize)
meanSkullCsize <- aggregate(Csize, 
                       by = list(pointLabels$Taxon), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL
rownames(meanSkullCsize) <- meanSkullCsize$Group.1
colnames(meanSkullCsize) <- c("Taxon", "Csize")
meanSkullCsize <- droplevels(meanSkullCsize)

#reorder Csize to fit the tree
meanSkullCsize <- meanSkullCsize[order(match(meanSkullCsize[,1],rawTree$tip.label)),]

#checks
rownames(meanSkullCsize) == rawTree$tip.label
dimnames(meanCoords)[[3]] == rawTree$tip.label
rownames(meanSkullCsize) == dimnames(meanCoords)[[3]]

allometry_gdf <- geomorph.data.frame(coords = meanCoords, Csize = meanSkullCsize$Csize, phy=rawTree)


allometry_fit <- procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdf)

par(mfrow=c(2,2))

summary(allometry_fit)

#Andrewsarchus is a high-leverage outlier
plot(allometry_fit)


```

#Allometry without Andrewsarchus - residuals are appropriately distributed
```{r}
#take out Andrewsarchus

rawTree <- drop.tip(phy = rawTree, tip = "Andrewsarchus_mongoliensis")

#shape

new.subset <-coords.subset(A = skullLmkSymm$symm.shape[,,-1],
                           group = droplevels( pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL

#reorder coords to fit the tree
meanCoords <- meanCoords[,,order(match(dimnames(meanCoords)[[3]],rawTree$tip.label))]



rawTree$tip.label == dimnames(meanCoords)[[3]]


#Csize without Andrewsarchus
Csize <- as.numeric(skullGPA$Csize)[-1]
meanSkullCsize <- aggregate(Csize, 
                       by = list(pointLabels$Taxon[-1]), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL

rownames(meanSkullCsize) <- meanSkullCsize$Group.1
colnames(meanSkullCsize) <- c("Taxon", "Csize")
meanSkullCsize <- droplevels(meanSkullCsize)

#reorder Csize to fit the tree
meanSkullCsize <- meanSkullCsize[order(match(meanSkullCsize[,1],rawTree$tip.label)),]

#checks
rownames(meanSkullCsize) == rawTree$tip.label
dimnames(meanCoords)[[3]] == rawTree$tip.label
rownames(meanSkullCsize) == dimnames(meanCoords)[[3]]


allometry_gdf <- geomorph.data.frame(coords = meanCoords, Csize = meanSkullCsize$Csize, phy=rawTree)


allometry_fit <- procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdf)

summary(allometry_fit)

#Here, the residuals are more appropriate
plot(allometry_fit)

Allom_PCA <- gm.prcomp(allometry_gdf$coords)


PCA_scores <-  as.data.frame(Allom_PCA$x)
Csize_taxa <- data.frame(Csize=meanSkullCsize$Csize, species = meanSkullCsize$Taxon)

#Check
Csize_taxa$species ==rownames(Allom_PCA$x)

Data_for_pgls <- cbind (PCA_scores, Csize_taxa)

Csize_PCA_frame <- comparative.data( rawTree, Data_for_pgls, names.col = species)

test <- pgls(Comp1~ log(Csize), data=Csize_PCA_frame)


summary(test)



```


#We are justified in keeping Andrewsarchus in the PCA plots because it does not impact on the morphospaces much. The below recreates Fig. 1 without Andrewsarchus.


```{r Labels}
pointLabels$PtCol <- "#fff0e0"
pointLabels$PtCol[pointLabels$Clade == "Metatheria"] <- "#e0f0ff"
pointLabels$PtCol[pointLabels$Family == "Thylacinidae"] <- "#c87ef8" #purple
pointLabels$PtCol[pointLabels$Taxon == "Borhyaena_tuberata"] <- "#ca576c" #red - just to see position. 
pointLabels$PtCol[pointLabels$Family == "Canidae"] <- "#ffaa5c" #blue
pointLabels$PtCol[pointLabels$Taxon == "Canis_lupus"] <- "#233045" #navy - just to see position
pointLabels$PtCol[pointLabels$Taxon == "Andrewsarchus_mongoliensis"] <- "#72c496" #green - just to see position
pointLabels$Point <- 21
pointLabels$Point[pointLabels$Clade == "Metatheria"] <- 23
pointLabels$ID <- str_c(substr(pointLabels$Taxon,1,3), str_to_title(substr(sub(".*_", "", pointLabels$Taxon),1,3)) )
# For publication, we'll change those 'position' colours back to their clade colours and use text call-outs


# Gather mean value labels
meanLabels <- pointLabels
# This pulls out labels by species, and keeps the rest of the associated data columns
meanLabels <- meanLabels %>%
    distinct(Taxon, .keep_all = T)
row.names(meanLabels) <- meanLabels$Taxon
meanLabels <- meanLabels[order(meanLabels$Taxon, decreasing = F), ]
```

#### Skull without Andrewsarchus
```{r Skull Set}

new.subset <-coords.subset(A = skullLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



skullPCA <- gm.prcomp(meanCoords) 
skullPCPlot <- plot(skullPCA)
PCScores <- skullPCA$x[,1:10]
skullPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```

#### Rostrum without Andrewsarchus
```{r Skull Set}
new.subset <-coords.subset(A = rostralLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



rostPCA <- gm.prcomp(meanCoords) 
rostPCPlot <- plot(rostPCA)
PCScores <- rostPCA$x[,1:10]
rostPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```
#### Neurocranium without Andrewsarchus
```{r Skull Set}
new.subset <-coords.subset(A = neuroLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



neurPCA <- gm.prcomp(meanCoords) 
neurPCPlot <- plot(neurPCA)
PCScores <- neurPCA$x[,1:10]
neurPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```

#Note that PC2 is flipped but otherwise very similar to Fig. 1
```{r Plot}

pdf(file="./Figures/Supp1_AMong_removed.pdf", width=10, height=10)

par(mfrow=c(2,2))

plot(NULL,
     xlim = range(skullPlotData$Comp1),
     ylim = -range(skullPlotData$Comp2),
     lwd = 0.5,
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Full Cranium")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(skullPlotData$Comp2 ~ skullPlotData$Comp1, 
     pch = skullPlotData$Point, 
     col = "#000000", 
     bg = skullPlotData$PtCol, 
     cex = 2)


#rostrum

plot(NULL,
     xlim = range(rostPlotData$Comp1),
     ylim = -range(rostPlotData$Comp2),
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Rostrum")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(rostPlotData$Comp2 ~ rostPlotData$Comp1, 
     pch = rostPlotData$Point, 
     col = "#000000", 
     bg = rostPlotData$PtCol, 
     cex = 2)



#Neurocranium

plot(NULL,
     xlim = range(neurPlotData$Comp1),
     ylim = -range(neurPlotData$Comp2),
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Neurocranium")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(neurPlotData$Comp2 ~ neurPlotData$Comp1, 
     pch = neurPlotData$Point, 
     col = "#000000", 
     bg = neurPlotData$PtCol, 
     cex = 2)

legend("bottomleft",
       legend = expression("Placentals", "Canids",
                  "Metatherians", "Thylacinidae", italic("Borhyaena"), italic("C. lupus")),
       pch = c(21, 21, 23, 23),
       col = c("#000000", "#000000", "#000000", "#000000"),
       pt.bg = c("#fff0e0", "#ffaa5c", "#e0f0ff", "#c87ef8", "#ca576c", "black"),
       bty = "n",
       pt.cex = 2,
       cex = 1,
       text.col = "black",
       horiz = F ,
       inset = c(0.00, 0.0))

plot.new()



dev.off()

```


#Allometry of the various partitions - all without Andrewsarchus

```{r}

#take out Andrewsarchus

rawTree <- drop.tip(phy = rawTree, tip = "Andrewsarchus_mongoliensis")

#take out Andrewsarchus for all partitions

partition_list_raw <- list(skull=skullLmkSymm, rostral=rostralLmkSymm, neuro=neuroLmkSymm, fnp=fnpLmkSymm, pa=paLmkSymm, mes=mesLmkSymm)
partition_list_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

meancoords_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for(i in 1:length(partition_list_noAnd)){

partition_list_noAnd[[i]] <-coords.subset(A = partition_list_raw[[i]]$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meancoords_noAnd[[i]] <- sapply(partition_list_noAnd[[i]], 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL
#reorder coords to fit the tree
meancoords_noAnd[[i]] <- meancoords_noAnd[[i]][,,order(match(dimnames(meancoords_noAnd[[i]])[[3]],rawTree$tip.label))]

}

#Check
rawTree$tip.label == dimnames(meancoords_noAnd[[3]]) [[3]]

#Csizes

csizes_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)
meanSkullCsize_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for(i in 1:length(partition_list_noAnd)){
  
csizes_noAnd[[i]] <- as.numeric(get(paste(names(csizes_noAnd)[i], "GPA", sep=""))$Csize)[-1]  
 
meanSkullCsize_noAnd[[i]] <- aggregate(csizes_noAnd[[i]], 
                       by = list(pointLabels$Taxon[-1]), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL

rownames(meanSkullCsize_noAnd[[i]]) <- meanSkullCsize$Group.1
colnames(meanSkullCsize_noAnd[[i]]) <- c("Taxon", "Csize")
meanSkullCsize_noAnd[[i]] <- droplevels(meanSkullCsize_noAnd[[i]]) 
 
#rearrange to fit tree
meanSkullCsize_noAnd[[i]] <- meanSkullCsize_noAnd[[i]][order(match(meanSkullCsize_noAnd[[i]][,1],rawTree$tip.label)),] 
  
  
  
}

#Now for the allometry analyses

#quick check that all lines up
meanSkullCsize_noAnd[[1]]$Taxon==dimnames(meancoords_noAnd$skull)[[3]]
dimnames(meancoords_noAnd$skull)[[3]] == rawTree$tip.label

allometry_gdfs <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)
allometry_fits <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for (i in 1:length(partition_list_noAnd)){
  
allometry_gdfs[[i]] <- geomorph.data.frame(coords = meancoords_noAnd[[i]], Csize = meanSkullCsize_noAnd[[i]]$Csize, taxon = meanSkullCsize_noAnd[[i]]$Taxon, phy=rawTree)  

allometry_fits[[i]] <- summary(procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdfs[[i]]))

}

#make output table

allometry_output <- matrix(nrow=length(partition_list_noAnd), ncol=length(allometry_fits$skull$table[1,]))
rownames(allometry_output) <- names(partition_list_noAnd)
colnames(allometry_output) <- colnames(allometry_fits$skull$table[1,] )

for (i in 1:length(partition_list_noAnd)){
  
 allometry_output[i,]<- unlist(allometry_fits[[i]]$table[1,])
  
  
}

#just so the processed talbe doesn't overwrite the previous
allometry_output_table <- allometry_output

allometry_output_table[,c(2,3,5,6,7)] <- sprintf("%.3f", allometry_output[,c(2,3,5,6,7)] )
allometry_output_table[,c(4)] <- sprintf("%.2f", allometry_output[,c(4)])

allometry_output_table<- allometry_output_table[,-c(1,2)]

allometry_output_table 


write.csv(allometry_output_table, file = "./Figures/Allometry_output.csv")

```



#adding the PCA 

```{r}

Allom_PCAs <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)
 temp <-NULL

#Data frames with PC scores and Csize bound
 
for (i in 1:length(Allom_PCAs)){
  
temp<- gm.prcomp(allometry_gdfs[[i]]$coords)
Allom_PCAs[[i]]$PCA_Csize <- cbind(as.data.frame(temp$x), as.numeric(allometry_gdfs[[i]]$Csize), as.character(allometry_gdfs[[i]]$taxon))

#soft coding this, messy but safer in case the pc numbers change
colnames(Allom_PCAs[[i]]$PCA_Csize)[length(colnames(Allom_PCAs[[i]]$PCA_Csize))-1] <- "Csize"
colnames(Allom_PCAs[[i]]$PCA_Csize)[length(colnames(Allom_PCAs[[i]]$PCA_Csize))]<- "species"
Allom_PCAs[[i]]$phy <-allometry_gdfs[[i]]$phy 


  
}



#Check checking
rownames(Allom_PCAs$rostral$PCA_Csize) == Allom_PCAs$neuro$phy$tip.label
Allom_PCAs$rostral$PCA_Csize$species == Allom_PCAs$neuro$phy$tip.label
  

Csize_PCA_frames <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for(i in 1:length(Csize_PCA_frames)){

Csize_PCA_frames [[i]] <- comparative.data( rawTree , 
                                            Allom_PCAs[[i]]$PCA_Csize, 
                                            'species')

}


#Setting up output tables
PCs_tested <- 10
pgls_PCA_table <- matrix(nrow=PCs_tested, ncol = 5)
colnames(pgls_PCA_table) <- c("Estimate",  "Std. Error",   "t value", "Rsq", "p")
rownames(pgls_PCA_table)<- paste (rep("PC", PCs_tested), c(1:PCs_tested), sep="")


comparative_data_PCA_outputs <- list(skull=pgls_PCA_table, rostral=pgls_PCA_table, neuro=pgls_PCA_table, fnp=pgls_PCA_table, pa=pgls_PCA_table, mes=pgls_PCA_table)


for (i in 1: length(Csize_PCA_frames)){
  
  for (j in 1: PCs_tested){
  
    
  temp <- summary(pgls( get(paste("Comp", j, sep="")) ~ log(Csize), data=Csize_PCA_frames[[i]]))
  
  comparative_data_PCA_outputs[[i]] [j,1:3]<- sprintf("%.3f",temp$coefficients[2,1:3])
 comparative_data_PCA_outputs[[i]][j,4]<- sprintf("%.2f", temp$r.squared)
 comparative_data_PCA_outputs[[i]][j,5]<- sprintf("%.3f", temp$coefficients[2,4])
 
  }
  
}

  

comparative_data_PCA_outputs$neurocranium

#OK packaging all of this into one huge table. stringsAsFactors = FALSE makes it so not everything turns into a character

PCA_All_outputs <-data.frame( rbind(c("Cranium", rep(NA, ncol(comparative_data_PCA_outputs$skull)-2)), comparative_data_PCA_outputs$skull[,-1],
      c("Rostrum", rep(NA, ncol(comparative_data_PCA_outputs$rostral)-2)), comparative_data_PCA_outputs$rostral[,-1],
      c("Neurocranium", rep(NA, ncol(comparative_data_PCA_outputs$neuro)-2)), comparative_data_PCA_outputs$neuro[,-1],
      c("Frontonasal Process", rep(NA, ncol(comparative_data_PCA_outputs$fnp)-2)), comparative_data_PCA_outputs$fnp[,-1],
      c("First Pharyngeal Arch", rep(NA, ncol(comparative_data_PCA_outputs$pa)-2)), comparative_data_PCA_outputs$pa[,-1],
      c("Paraxial Mesoderm", rep(NA, ncol(comparative_data_PCA_outputs$mes)-2)), comparative_data_PCA_outputs$mes[,-1]
),stringsAsFactors = FALSE)



write.csv(PCA_All_outputs, file="./Figures/pgls_PCA_tables.csv")



#because this is a lot of coding and potential for human error, I'm here running the PCAs work - it checks out.

# plot(Csize_PCA_frames$skull$data$Comp2~Csize_PCA_frames$skull$data$Comp1, asp=TRUE)
# text(Csize_PCA_frames$skull$data$Comp2~Csize_PCA_frames$skull$data$Comp1, asp=TRUE, labels=rownames(Csize_PCA_frames$skull$data) )
# 

```


#Making Table 1 - allometry
```{r}

allometry_output_table

PC12_list <- matrix(nrow=length(comparative_data_PCA_outputs), ncol=4)

for (i in 1:length(comparative_data_PCA_outputs)){
  
PC12_list[i,] <- c (comparative_data_PCA_outputs[[i]][c(1), c(4,5)], comparative_data_PCA_outputs[[i]][c(2), c(4,5)])
}


Table_1 <- cbind(allometry_output_table, PC12_list)

colnames(Table_1)[c(5:9)]<- c("p", "PC1_Rsq", "PC1_p", "PC2_Rsq", "PC2_p")


Table_1 <- data.frame(Table_1, stringsAsFactors = FALSE)

write.csv(Table_1, file="./Figures/Table_1_Allometry.csv")

```

#Allometry of separate clades,the most appropriate for this dataset are marsupials, feliforms, and canids. 

#create allometry ANCOVAs with clades as factor. 

```{r}

#We already have all partitions - these are allometry_gdfs. Now we have to split these into allometry_gdfs for each clade. Starting off with just one partition

#taxa match when removing Andrewsarchus from mean labels, but need to be reordered. Also Andrewsarchus needs removing as a level after removal.
meanLabels_noAnd <-meanLabels[-2,]
meanLabels_noAnd$Taxon <- droplevels (meanLabels_noAnd$Taxon)
allometry_gdfs$skull$taxon %in% meanLabels_noAnd[-2,]$Taxon
allometry_gdfs$skull$taxon == meanLabels_noAnd$Taxon

#So we now reorder the labels according to the gdf labels
Labels_allometry_partitions <- meanLabels_noAnd[match(allometry_gdfs$skull$taxon,meanLabels_noAnd$Taxon),]

#Now the labels are the right order and gdf taxon order remains unchanged and the trees in the gdfs also talks to the labels
allometry_gdfs$skull$taxon==Labels_allometry_partitions$Taxon

allometry_gdfs$skull$phy$tip.label == Labels_allometry_partitions$Taxon

#add clade information to gdfs and drop whippomorphan (Artiodactyla) level
allometry_clade_gdfs <- allometry_gdfs
allometry_clade_fits <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, Fnp=NULL, Pa=NULL, Mes=NULL)

for (i in 1:length( allometry_gdfs)){

  
  allometry_clade_gdfs[[i]]$labels <- Labels_allometry_partitions
  allometry_clade_gdfs[[i]]$labels$Clade <- droplevels(allometry_clade_gdfs[[i]]$labels$Clade)
  names(allometry_clade_gdfs[[i]]$Csize) <- dimnames(allometry_clade_gdfs[[i]]$coords)[[3]]
}



#all 59, as it should be, and all data contained in here
allometry_clade_gdfs[[i]]$phy
dimnames(allometry_clade_gdfs[[i]]$coords)[[3]]
allometry_clade_gdfs[[i]]$Csize
nrow(allometry_clade_gdfs[[i]]$labels)
names(allometry_clade_gdfs)

allometry_clade_gdfs$skull$labels

#sample sizes:

length(which(Labels_allometry_partitions$Clade=="Feliformia"))
length(which(Labels_allometry_partitions$Clade=="Caniformia"))
length(which(Labels_allometry_partitions$Clade=="Metatheria"))

rawtree_clades <- rawTree

#If you want, here you can take borhyaena from the tree, make sure it talks to the gdf etc
#drop.tip (rawTree, which(as.vector(Labels_allometry_partitions$Family) == "Borhyaenidae")) 
#rawtree_clades$tip.label == dimnames(allometry_clade_gdfs[[1]]$coords)[[3]]
#plot(rawtree_clades)

#Tests below- always check for small mistakes that throw everything. For reasons I do NOT understand (grr!) I can't use the data= at the end and have to spell out the whole address. Doesn't really matter but I could swear that this didn't use to work and it had to be the other way!

#testfit <- summary(procD.pgls(allometry_clade_gdfs$skull$coords ~ log(allometry_clade_gdfs$skull$Csize), phy=rawtree_clades))

#just checking that the thylacine is still one of the largest-skulled species in the sample: 

#allometry_clade_gdfs$skull$Csize[order(log(allometry_clade_gdfs$skull$Csize))]

#also making sure the shape still is associated with the right species - all good, hashing out
#spheres3d( allometry_clade_gdfs$skull$coords[,,which (dimnames(allometry_clade_gdfs$skull$coords)[[3]]=="Thylacinus_cynocephalus")], radius = 0.001, col="hotpink")

#spheres3d( allometry_clade_gdfs$rostral$coords[,,which (dimnames(allometry_clade_gdfs$rostral$coords)[[3]]=="Thylacinus_cynocephalus")], radius = 0.001, col="hotpink")

```

# moving on to by-clade analysis, we were initially not sure if pgls or lm are better here (because below we use lm to compare clades), we ended up settling for pgls but the point is the same for both- different partitions have different allometries in different clades. 
#first checking if there even is allometry within the clades. 
```{r}


#the coordinates and csize could all be subsectioned into the main clades within the loop, but that makes it unintelligible. So the loop below creates a list of vectors that pulls out the species within each clade outside the loop.


#Create a list for allometries for individual clades. I have this from co-pilot, nice quick workflow!
allom_gdf_by_clade <- setNames(vector("list", 3), paste(levels(allometry_clade_gdfs$skull$labels$Clade)))


for (i in 1: length(levels(allometry_clade_gdfs$skull$labels$Clade))){
  
  allom_gdf_by_clade[[i]] <- which(allometry_clade_gdfs$skull$labels$Clade==levels(allometry_clade_gdfs$skull$labels$Clade)[i])
  
}


#Create a list with the 6 partitions and each partition has the 3 species - again, Copilot is a gem here. Use "replicate" to make lists with multiple named sublists
allometry_clade_alloms_within_clade <- replicate(length(allometry_clade_gdfs), 
                                                 setNames(vector("list",
                                                          length(levels(allometry_clade_gdfs$skull$labels$Clade))),
                                                          paste(levels(allometry_clade_gdfs$skull$labels$Clade))),
                                                          simplify=FALSE
                                                 )
#then name the list elements one level up
names(allometry_clade_alloms_within_clade) <- names(allometry_clade_gdfs)

#Below for lm

for (i in 1: length(allometry_clade_gdfs)){

  for (j in 1:length(levels(allometry_clade_gdfs$skull$labels$Clade))){

  allometry_clade_alloms_within_clade[[i]][[j]] <- summary(
                                                  procD.lm(
      allometry_clade_gdfs[[i]]$coords[,,allom_gdf_by_clade[[j]]] ~        log(allometry_clade_gdfs[[i]]$Csize[allom_gdf_by_clade[[j]]] )
      )
      )

  }
}

#Then PGLS

allometry_clade_alloms_within_clade_pgls <- allometry_clade_alloms_within_clade

for (i in 1: length(allometry_clade_gdfs)){
  
  for (j in 1:length(levels(allometry_clade_gdfs$skull$labels$Clade))){
    
  allometry_clade_alloms_within_clade_pgls[[i]][[j]] <- summary(
                                                  procD.pgls(
      allometry_clade_gdfs[[i]]$coords[,,allom_gdf_by_clade[[j]]] ~        log(allometry_clade_gdfs[[i]]$Csize[allom_gdf_by_clade[[j]]] ),
          phy = keep.tip(allometry_clade_gdfs$skull$phy,
            names(allometry_clade_gdfs$skull$Csize[allom_gdf_by_clade[[j]]]))
      )
      )
      
  }
}

cani_test <- keep.tip(allometry_clade_gdfs$skull$phy, names(allometry_clade_gdfs$skull$Csize[allom_gdf_by_clade$Caniformia]))

cani_test$tip.label == names(allometry_clade_gdfs$skull$Csize[allom_gdf_by_clade$Caniformia])

#This seems to have worked, but just to check that it's really working - the below list vs manual coding show identical results

allometry_clade_alloms_within_clade$rostral$Caniformia

summary(procD.lm(allometry_clade_gdfs$rostral$coords[,,which(allometry_clade_gdfs$skull$labels$Clade== "Caniformia")] ~ log(allometry_clade_gdfs$rostral$Csize[which(allometry_clade_gdfs$skull$labels$Clade== "Caniformia") ] )))

#Now pull out the relevant values for a table 
```

#Table. I am here for space reasons relying on replacing allometry_clade_alloms_within_clade with allometry_clade_alloms_within_clade_pgls for generating teh table csv. Not elegant, but does the job
```{r}
#Matrix with four rows per comparison (partition name, size, clade and interaction)
Allometry_table_within_each_clade <- matrix (nrow=(length(allometry_clade_alloms_within_clade_pgls)*4), ncol = 4)
colnames(Allometry_table_within_each_clade)<-c("SS", "F", "RSq", "p" )

rownames(Allometry_table_within_each_clade)<- rep ( c(NA,levels(allometry_clade_gdfs$skull$labels$Clade)), length(allometry_clade_alloms_within_clade_pgls))
#Row names just for the partitions
rownames(Allometry_table_within_each_clade)[c(1,c(1:(length(names(allometry_clade_gdfs))-1)*4)+1)] <- names(allometry_clade_alloms_within_clade_pgls)

#The loop. Surely there is an easier way!

for (i in 1: length (allometry_clade_fits)){
  
   for (j in 1: length(levels(allometry_clade_gdfs$skull$labels$Clade))){

  #SS
  Allometry_table_within_each_clade[(((i)*4)+j-3),1] <- round(allometry_clade_alloms_within_clade_pgls[[i]][[j]]$table$SS[1], 3)
  #Z - I don't see the point of two similar metrics of effect size, hopefully I'm not wrong but I don't think it is informative
  #Allometry_table_within_each_clade[(((i)*4)+j-3),2] <- round(allometry_clade_alloms_within_clade_pgls[[i]][[j]]$table$Z[1], 3)
  #F
  Allometry_table_within_each_clade[(((i)*4)+j-3),2] <- round(allometry_clade_alloms_within_clade_pgls[[i]][[j]]$table$F[1], 3)
  #Rsq
  Allometry_table_within_each_clade[(((i)*4)+j-3),3] <- round(allometry_clade_alloms_within_clade_pgls[[i]][[j]]$table$Rsq[1],2)
  #p
  Allometry_table_within_each_clade[(((i)*4)+j-3),4] <- round(allometry_clade_alloms_within_clade_pgls[[i]][[j]]$table$`Pr(>F)`[1], 3)
  
  }
  
} 
 

write.csv(Allometry_table_within_each_clade, file="./Figures/Allometry_table_within_each_clade.csv") 
#check check
summary(procD.lm(allometry_clade_gdfs$mes$coords[,,which(allometry_clade_gdfs$mes$labels$Clade== "Metatheria")] ~ log(allometry_clade_gdfs$mes$Csize[which(allometry_clade_gdfs$mes$labels$Clade== "Metatheria") ] )))


```

Now we also want to visualise the shape variation within each clade
```{r}

#First, create by-clade allom plots for the allom-by-clade fit.For this, we need both the fits and the plots.

#Unfortunately, we used the summaries not the fits above, so now we have to make the fits first. Here first over-writing the allometry_clade_alloms_within_clade object containing the summaries.

allometry_clade_alloms_within_clade_fits <- allometry_clade_alloms_within_clade

for (i in 1: length(allometry_clade_gdfs)){
  
  for (j in 1:length(levels(allometry_clade_gdfs$skull$labels$Clade))){
    
  allometry_clade_alloms_within_clade_fits[[i]][[j]] <- 
                                                  procD.lm(
      allometry_clade_gdfs[[i]]$coords[,,allom_gdf_by_clade[[j]]] ~        log(allometry_clade_gdfs[[i]]$Csize[allom_gdf_by_clade[[j]]] )
      )
      
      
  }
}


#Making a plot list with the right names, then run code
Allom_plots_within_each_clade <- allometry_clade_alloms_within_clade

for (i in 1: length (allometry_clade_alloms_within_clade)){
  
   for (j in 1: length(allometry_clade_alloms_within_clade_fits$skull)){

Allom_plots_within_each_clade [[i]][[j]] <- plot(allometry_clade_alloms_within_clade_fits[[i]][[j]],
                                            type="regression", 
                                            #This needs a wordy split of the gdf dataset by subgroup, how annoying
                                            predictor = log(allometry_clade_gdfs[[i]]$Csize[allom_gdf_by_clade[[j]]] ),
                                            reg.type = "RegScore")

   }
}

#Now making another loop for predicted shapes

#Making a plot list with the right names, then run code
Allom_preds <- allometry_clade_alloms_within_clade

for (i in 1: length (allometry_clade_alloms_within_clade)){
  
   for (j in 1: length(allometry_clade_alloms_within_clade_fits$skull)){

Allom_preds [[i]][[j]] <- shape.predictor(allometry_clade_alloms_within_clade_fits[[i]][[j]]$GM$fitted,
                                          x=Allom_plots_within_each_clade[[i]][[j]]$RegScore,
                                          predmin=min(Allom_plots_within_each_clade[[i]][[j]]$RegScore), 
                                          predmax=max(Allom_plots_within_each_clade[[i]][[j]]$RegScore)
                                          )
   }
}
```

Producing the displacement graphs could be done in a loop and usrmats but that would require tons of fiddling with imageMagick etc. to remove the white space. So here it is more manually with screenshots saved (and I had to jiggle the radii for PA (0.0001) and MES (0.001) because they became tiny on the plot)

```{r}

predcols <- c("#fff0e0", "#ffaa5c", "#e0f0ff")



for (i in 1: length (allometry_clade_alloms_within_clade)){

#ensures that a new window is opened  
open3d()  
mfrow3d(1,3,sharedMouse = TRUE, byrow = TRUE)
 
  
   for (j in 1: length(allometry_clade_alloms_within_clade_fits$skull)){

#set up
panel_w <- 700
panel_h <- 700
border  <- 0  # tweak if your window manager adds borders you want to compensate
par3d(windowRect = c(100, 100, 100 + 3*panel_w + 2*border, 100 + panel_h + 2*border))

par3d(skipRedraw=TRUE)

#next3d() needs to be first, not last, otherwise the first window goes blank.
next3d()

#making colourised spheres according to the individual allometry plots in the supps
spheres3d(Allom_preds[[i]][[j]]$predmin, col=predcols[j], radius=0.0015)

#plotreftotarget with slightly smaller radii - unfortunately, the colour can't easily be added to plotreftotarget
plotRefToTarget(Allom_preds[[i]][[j]]$predmin, Allom_preds[[i]][[j]]$predmax, method="vector", radius=0.0009, add=TRUE)

par3d(skipRedraw = FALSE)

   }

}
                
```

Specifically asking if small-thylacine displacement resembles small-large displacement - not published but we were curious after a reviewer asked

```{r}

#ensures that a new window is opened  
open3d()  
mfrow3d(1,2,sharedMouse = TRUE)

plotRefToTarget(Allom_preds$skull$Metatheria$predmin, Allom_preds$skull$Metatheria$predmax, method="vector" )

#this is allometry towards largest, but it includes thylacine and sarcophilus as high-leverage outliers
plotRefToTarget(Allom_preds$skull$Metatheria$predmin, allometry_gdfs$skull$coords[,,which(dimnames(allometry_gdfs$skull$coords)[[3]]=="Thylacinus_cynocephalus")], method="vector")

#So now taking out those two to ask if the thylacine is a scaled-up smaller marsupial carnivore and not considering any of the large high-leverage outliers - thylacine, devil, borhyaena. But no, at this level, allometry is not even significant any more - small samples size is an issue here for sure but also there's just NO STRONG EFFECT

#Se[arate out marsupials first so the code doesn't get too long
Skull_coords_mars <-allometry_gdfs$skull$coords[,,allom_gdf_by_clade$Metatheria]

Skull_coords_mars_noTS <- Skull_coords_mars [,,-(which(dimnames(Skull_coords_mars)[[3]]%in% c( "Borhyaena_tuberata", "Thylacinus_cynocephalus", "Sarcophilus_harrisii")))]


Skull_size_mars <- allometry_gdfs$skull$Csize[allom_gdf_by_clade$Metatheria]
names(Skull_size_mars) <- dimnames(Skull_coords_mars)[[3]]

Skull_size_mars_noTS <- Skull_size_mars [-(which(dimnames(Skull_coords_mars)[[3]]%in% c( "Borhyaena_tuberata", "Thylacinus_cynocephalus", "Sarcophilus_harrisii")))]

tree_mars <- keep.tip(rawTree, names(Skull_size_mars_noTS))

match(tree_mars$tip.label,names(Skull_size_mars_noTS))

small_das_fit <- procD.pgls(Skull_coords_mars_noTS~ log(Skull_size_mars_noTS), tree_mars)

summary(small_das_fit)


```


```{r}

#Caniforms, feliforms, marsupials;  labels$Clade . Note that we had previously run this with Borhyaena removed (it is a non-marsupial so we were on the fence about including it), but the results are nearly the same with slightly smaller (0.01 max different) Rsquareds



for (i in 1: length(allometry_clade_gdfs)){
  
  allometry_clade_fits[[i]] <- summary(procD.lm(allometry_clade_gdfs[[i]]$coords ~ log(allometry_clade_gdfs[[i]]$Csize)* allometry_clade_gdfs[[i]]$labels$Clade))
      
}


#Matrix with four rows per comparison (partition name, size, clade and interaction)
Allometry_table_clade <- matrix (nrow=(length(allometry_clade_fits)*4), ncol = 5)

#the below code needs this line to start
rownames(Allometry_table_clade)<- rep (names(allometry_clade_fits), each=4)

for (i in 2: length(allometry_clade_fits)){
#need to start off with one  
rownames(Allometry_table_clade)[1:4] <- c(names(allometry_clade_fits)[1], "log Csize", "Clade", "Clade:Csize")
#then add to the previous - there must be a nicer way of coding this, but I haven't found it and this is still better than repeated pastes
rownames(Allometry_table_clade)[(((i-1)*4)+1):(i*4)]<- c(names(allometry_clade_fits)[i], "log Csize", "Clade", "Clade:Csize")

}

colnames(Allometry_table_clade)<-c("SS", "RSq", "F", "Z", "p" )

#Now fill in the results

for (i in 2: length (allometry_clade_fits)){
  
#again, 1 needs separate treatment
  #SS
  Allometry_table_clade[2:4,1] <- round(allometry_clade_fits[[1]]$table$SS[1:3], 3)
  #Rsq
  Allometry_table_clade[2:4,2] <- round(allometry_clade_fits[[1]]$table$Rsq[1:3], 2)
  #F
  Allometry_table_clade[2:4,3] <- round(allometry_clade_fits[[1]]$table$F[1:3], 3)
  #F
  Allometry_table_clade[2:4,4] <- round(allometry_clade_fits[[1]]$table$Z[1:3], 3)
  #p
  Allometry_table_clade[2:4,5] <- round(allometry_clade_fits[[1]]$table$`Pr(>F)`[1:3], 3)
  
  #now all the others. i-1*4+2 here (not +1) because the row with the name of teh partition is blank
  #SS
  Allometry_table_clade[(((i-1)*4)+2):(i*4),1] <- round(allometry_clade_fits[[i]]$table$SS[1:3], 3)
  #Rsq
  Allometry_table_clade[(((i-1)*4)+2):(i*4),2] <- round(allometry_clade_fits[[i]]$table$Rsq[1:3], 2)
  #F
  Allometry_table_clade[(((i-1)*4)+2):(i*4),3] <- round(allometry_clade_fits[[i]]$table$F[1:3], 3)
  #F
  Allometry_table_clade[(((i-1)*4)+2):(i*4),4] <- round(allometry_clade_fits[[i]]$table$Z[1:3], 3)
  #p
  Allometry_table_clade[(((i-1)*4)+2):(i*4),5] <- round(allometry_clade_fits[[i]]$table$`Pr(>F)`[1:3], 3)
  
  }


write.csv(Allometry_table_clade, file="./Figures/Allometry_table.csv")
```

#Regscore plots - here using the fit objects to contain the fits, not the summaries

#NOTE there is no point using clade as covariate if using pgls, this guarantees a non-significant result because the effect your're looking for is explicitly washed out of the data. But for some reason you feel forced to do pgls or want to play with a different factor, change to procD.pgls function and add "phy=rawtree_clades"

```{r}

#This is a plot of overall allometry without accounting for differences among clades


for (i in 1: length(allometry_clade_gdfs)){
  
  allometry_clade_fits[[i]] <- procD.lm(allometry_clade_gdfs[[i]]$coords ~ log(allometry_clade_gdfs[[i]]$Csize))
      
}

#Making a plot list with the right names
Allom_plots <- allometry_clade_gdfs

for (i in 1: length(allometry_clade_gdfs)){
  
  Allom_plots[[i]]<- plot(allometry_clade_fits[[i]], type="regression", predictor = log(allometry_clade_gdfs[[i]]$Csize))
  
}


```

#Full supplementary Figure plots

```{r}
#Set up colours

skullPlotData_Allom <- skullPlotData

skullPlotData_Allom$PtCol[which(skullPlotData_Allom$Family=="Borhyaenidae")]<-"#e0f0ff"

PlotData_Allom <-skullPlotData_Allom[ match (names(allometry_clade_gdfs$skull$Csize),rownames(skullPlotData_Allom) ),]

pdf(file="./Figures/SuppX_Allometry_partitions.pdf", width=15, height=8)

par(mfrow=c(2,3), mar=c(4,4,3,1), mgp=c(2,0.5,0))

for (i in 1: length(allometry_clade_gdfs)){
  
  plot(Allom_plots[[i]]$RegScore~log(allometry_clade_gdfs[[i]]$Csize),
      xlab= "log Centroid Size of Partition",
      ylab= "Regression Score of Shape ~ log (Csize) Fit",
      main =  names(allometry_clade_fits)[i],
      cex.lab=1.2,
      cex.axis=1.2
         
         )
  
  points(Allom_plots[[i]]$RegScore~log(allometry_clade_gdfs[[i]]$Csize), 
     pch = PlotData_Allom$Point, 
     col = "#000000", 
     bg = PlotData_Allom$PtCol, 
     cex = 2)
  
 

text(Allom_plots[[i]]$RegScore~log(allometry_clade_gdfs[[i]]$Csize), labels = PlotData_Allom$ID)

  #abline(coef(lm(Allom_plots[[i]]$RegScore~log(allometry_clade_gdfs[[i]]$Csize))))

legend("topleft",
       legend = expression("Feliforms", "Caniforms",
                  "Marsupials", "Thylacinidae", italic("C. lupus")),
       pch = c(21, 21, 23, 23),
       col = c("#000000", "#000000", "#000000", "#000000"),
       pt.bg = c("#fff0e0", "#ffaa5c", "#e0f0ff", "#c87ef8", "black"),
       bty = "n",
       pt.cex = 2,
       cex = 1.2,
       text.col = "black",
       horiz = F ,
       inset = c(0.00, 0.0))

}

dev.off()

```


#slope comparison using pairwise
```{r}
#starting with one analysis

#so here I have to re-run the fit because the previous fit objects contained summaries.

#Running pairwise fit for SLOPE differences - relevant for all but neurocranium and MES
allometry_clade_SLOPE_fits <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, Fnp=NULL, Pa=NULL, Mes=NULL)

for (i in 1: length(allometry_clade_gdfs)){
  
  allometry_clade_SLOPE_fits[[i]] <- procD.lm(allometry_clade_gdfs[[i]]$coords ~ log(allometry_clade_gdfs[[i]]$Csize)* allometry_clade_gdfs[[i]]$labels$Clade)
      
}

#Running pairwise fit for SLOPE differences - relevant for all but neurocranium and MES
allometry_clade_MEAN_fits <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, Fnp=NULL, Pa=NULL, Mes=NULL)

for (i in 1: length(allometry_clade_gdfs)){
  
  allometry_clade_MEAN_fits[[i]] <- procD.lm(allometry_clade_gdfs[[i]]$coords ~ log(allometry_clade_gdfs[[i]]$Csize)+ allometry_clade_gdfs[[i]]$labels$Clade)
      
}

#test with just one - works
pairwisecheck <- pairwise(allometry_clade_SLOPE_fits$Cranium, groups=allometry_clade_gdfs[[i]]$labels$Clade, covariate =log(allometry_clade_gdfs$skull$Csize) )
summary (pairwisecheck)

summary.pairwise(pairwisecheck, test.type = "VC", angle.type = "deg")

#now the whole thing, going straight for summaries
#these are always the same, so coding them here
groups <- allometry_clade_gdfs$skull$labels$Clade

pairwise_tests_slope <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, FNP=NULL, PA=NULL, MES=NULL)

for (i in 1: length(allometry_clade_SLOPE_fits) ){
  
  pairwise_tests_slope[[i]] <- summary(pairwise(allometry_clade_SLOPE_fits[[i]], groups=groups, covariate =log(allometry_clade_gdfs[[i]]$Csize) ))

}

pairwise_tests_mean <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, FNP=NULL, PA=NULL, MES=NULL)


for (i in 1: length(allometry_clade_MEAN_fits) ){
  
  pairwise_tests_mean[[i]] <- summary(pairwise(allometry_clade_MEAN_fits[[i]], groups=groups))

}

#make table - this is the same for slope and mean results

#Matrix with four rows per comparison (partition name, size, clade and interaction)
pairwise_table_mean <- matrix (nrow=(length(pairwise_tests_mean)*4), ncol = 4)

# Make row names - the below code somehow needs this line to start
rownames(pairwise_table_mean)<- rep (names(pairwise_tests_mean), each=4)

for (i in 2: length(pairwise_tests_mean)){
#need to start off with one  
rownames(pairwise_table_mean)[1:4] <- c(names(pairwise_tests_mean)[[1]], rownames(pairwise_tests_mean[[1]]$summary.table))
#then add to the previous - there must be a nicer way of coding this, but I haven't found it and this is still better than repeated pastes
rownames(pairwise_table_mean)[(((i-1)*4)+1):(i*4)]<- c(names(pairwise_tests_mean)[[i]], rownames(pairwise_tests_mean[[i]]$summary.table))

}

colnames(pairwise_table_mean)<-c("d", "UCL 95%", "Z", "p_mean" )

#Slope table is identical, creating it here
pairwise_table_slope <- pairwise_table_mean
colnames(pairwise_table_slope)<-c("d", "UCL 95%", "Z", "p_slope" )

#Now fill in the results, running everything here becasue it's easier and doesn't really matter
#slopes first

for (i in 2: length (pairwise_tests_slope)){
  
#again, 1 needs separate treatment
  #d
  pairwise_table_slope[2:4,1] <- round(pairwise_tests_slope[[1]]$summary.table$d, 3)
  #UCL
  pairwise_table_slope[2:4,2] <- round(pairwise_tests_slope[[1]]$summary.table$`UCL (95%)`, 3)
  #Z
  pairwise_table_slope[2:4,3] <- round(pairwise_tests_slope[[1]]$summary.table$Z, 3)
  #p
  pairwise_table_slope[2:4,4] <- round(pairwise_tests_slope[[1]]$summary.table$`Pr > d`, 3)

  
  #now all the others. i-1*4+2 here (not +1) because the row with the name of the partition is blank
  #SS
  pairwise_table_slope[(((i-1)*4)+2):(i*4),1] <- round(pairwise_tests_slope[[i]]$summary.table$d, 3)
  #Rsq
  pairwise_table_slope[(((i-1)*4)+2):(i*4),2] <- round(pairwise_tests_slope[[i]]$summary.table$`UCL (95%)`, 3)
  #F
  pairwise_table_slope[(((i-1)*4)+2):(i*4),3] <- round(pairwise_tests_slope[[i]]$summary.table$Z, 3)
  #F
  pairwise_table_slope[(((i-1)*4)+2):(i*4),4] <- round(pairwise_tests_slope[[i]]$summary.table$`Pr > d`, 3)
  
}

#now mean
for (i in 2: length (pairwise_tests_mean)){
  
#again, 1 needs separate treatment
  #d
  pairwise_table_mean[2:4,1] <- round(pairwise_tests_mean[[1]]$summary.table$d, 3)
  #UCL
  pairwise_table_mean[2:4,2] <- round(pairwise_tests_mean[[1]]$summary.table$`UCL (95%)`, 3)
  #Z
  pairwise_table_mean[2:4,3] <- round(pairwise_tests_mean[[1]]$summary.table$Z,3)
  #p
  pairwise_table_mean[2:4,4] <- round(pairwise_tests_mean[[1]]$summary.table$`Pr > d`,3)

  
  #now all the others. i-1*4+2 here (not +1) because the row with the name of the partition is blank
  #SS
  pairwise_table_mean[(((i-1)*4)+2):(i*4),1] <- round(pairwise_tests_mean[[i]]$summary.table$d, 3)
  #Rsq
  pairwise_table_mean[(((i-1)*4)+2):(i*4),2] <- round(pairwise_tests_mean[[i]]$summary.table$`UCL (95%)`, 3)
  #F
  pairwise_table_mean[(((i-1)*4)+2):(i*4),3] <- round(pairwise_tests_mean[[i]]$summary.table$Z, 3)
  #F
  pairwise_table_mean[(((i-1)*4)+2):(i*4),4] <- round(pairwise_tests_mean[[i]]$summary.table$`Pr > d`, 3)
  
}




```

Now we also want to include the angles between slopes where we have them. Unfortunately, that requires a slightly different approach of 1) saving the pairwise tests and 2) running the pairwise.text functionality to specify that we're after the angles. Otherwise, the statistics are the same though. So just to get this over the line, doing another big bunch of coding here to append to the pairwise table slopes.

```{r}

pairwise_tests_slope_angles_pairwise_objects <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, FNP=NULL, PA=NULL, MES=NULL)

for (i in 1: length(allometry_clade_SLOPE_fits) ){
  
  pairwise_tests_slope_angles_pairwise_objects[[i]] <- pairwise(allometry_clade_SLOPE_fits[[i]], groups=groups, covariate =log(allometry_clade_gdfs[[i]]$Csize) )

}

pairwise_tests_slope_angles_pairwise_tests <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, FNP=NULL, PA=NULL, MES=NULL)

for (i in 1: length(pairwise_tests_slope_angles_pairwise_objects) ){
  
  pairwise_tests_slope_angles_pairwise_tests[[i]] <- summary.pairwise(pairwise_tests_slope_angles_pairwise_objects[[i]], test.type = "VC", angle.type = "deg")

}

#make table
pairwise_table_angles <- matrix (nrow=(length(pairwise_tests_slope_angles_pairwise_tests)*4), ncol = 5)
#Allometry_table_within_each_clade <- matrix (nrow=(length(allometry_clade_alloms_within_clade)*4), ncol = 4)
colnames(pairwise_table_angles)<-c("r", "angle", "UCL(95%)", "Z", "p_angles" )

rownames(pairwise_table_angles)<- rep ( c(NA,levels(allometry_clade_gdfs$skull$labels$Clade)), length(allometry_clade_alloms_within_clade))
#Row names just for the partitions
rownames(pairwise_table_angles)[c(1,c(1:(length(names(allometry_clade_gdfs))-1)*4)+1)] <- names(allometry_clade_alloms_within_clade)

#the loop

#now mean
for (i in 2: length (pairwise_tests_slope_angles_pairwise_tests)){
  
#again, 1 needs separate treatment
  #r
  pairwise_table_angles[2:4,1] <- round(pairwise_tests_slope_angles_pairwise_tests[[1]]$summary.table$r, 3)
  #angle
  pairwise_table_angles[2:4,2] <- round(pairwise_tests_slope_angles_pairwise_tests[[1]]$summary.table$angle, 1)
  #UCL
  pairwise_table_angles[2:4,3] <- round(pairwise_tests_slope_angles_pairwise_tests[[1]]$summary.table$`UCL (95%)`, 3)
  #Z
  pairwise_table_angles[2:4,4] <- round(pairwise_tests_slope_angles_pairwise_tests[[1]]$summary.table$Z, 3)
  #p
  pairwise_table_angles[2:4,5] <- round(pairwise_tests_slope_angles_pairwise_tests[[1]]$summary.table$`Pr > angle`, 3)
  
  
  #now the rest
    #r
   pairwise_table_angles[(((i-1)*4)+2):(i*4),1] <- round(pairwise_tests_slope_angles_pairwise_tests[[i]]$summary.table$r, 3)
  #angle
  pairwise_table_angles[(((i-1)*4)+2):(i*4),2] <- round(pairwise_tests_slope_angles_pairwise_tests[[i]]$summary.table$angle, 1)
  #UCL
  pairwise_table_angles[(((i-1)*4)+2):(i*4),3] <- round(pairwise_tests_slope_angles_pairwise_tests[[i]]$summary.table$`UCL (95%)`, 3)
  #Z
  pairwise_table_angles[(((i-1)*4)+2):(i*4),4] <- round(pairwise_tests_slope_angles_pairwise_tests[[i]]$summary.table$Z, 3)
  #p
  pairwise_table_angles[(((i-1)*4)+2):(i*4),5] <- round(pairwise_tests_slope_angles_pairwise_tests[[i]]$summary.table$`Pr > angle`, 3)
  }


write.csv(cbind (pairwise_table_slope,pairwise_table_angles,pairwise_table_mean), file="./Figures/Pairwise_tables.csv")

#lm.rrpp.fit does not usually adjust for multiple comparisons and the authors don't mention it. But the final editorial comments asked for information about adjustment and this made VW realise that there seems to be no good reason not to do them. To make it easier, we here adjust the p values with a simple benjamini hochberg adjustment after the fact
pairwise_table_bj<-(cbind (pairwise_table_slope,pairwise_table_angles,pairwise_table_mean))[,c(4,9,13)]
pairwise_table_p<-(cbind (pairwise_table_slope,pairwise_table_angles,pairwise_table_mean))[,c(4,9,13)]


#this loop just runs through all the p values


for (i in seq(1,nrow(pairwise_table_p)-3, by=4)){
  
  pairwise_table_bj[(i+1):(i+3),1]<-p.adjust(pairwise_table_p[(i+1):(i+3),1], method="hochberg")
  pairwise_table_bj[(i+1):(i+3),2]<-p.adjust(pairwise_table_p[(i+1):(i+3),2], method="hochberg")
  pairwise_table_bj[(i+1):(i+3),3]<-p.adjust(pairwise_table_p[(i+1):(i+3),3], method="hochberg")
  
  
}


write.csv(pairwise_table_bj, file = "Figures/Pairwise_adjusted_pvalues.csv")

#check check tick tick - always check! All good.set i to below 6  first.

pairwisecheck <- pairwise(allometry_clade_SLOPE_fits[[i]], groups=allometry_clade_gdfs[[i]]$labels$Clade, covariate =log(allometry_clade_gdfs$skull$Csize) )

#slopes
summary.pairwise(pairwisecheck, test.type = "dist")
#angles
summary.pairwise(pairwisecheck, test.type = "VC", angle.type = "deg")


```
