---
title: "01_01_Allometry_check"
author: "Vera Weisbecker"
date: "2024-05-03"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##We are ignoring the effect of allometry on shape because it is not very strong. 
```{r}

library(geomorph)  # CRAN v4.0.7
library(ape)       # CRAN v5.8
library(caper)     # CRAN v1.0.3
library(shapes)    # CRAN v1.2.7
library(tidyverse) # CRAN v2.0.0
library(svglite)   # CRAN .1.3
library(here)      # CRAN v1.0.1

load(here("Data", "Processed", "skullLmkSymm.rda"))
load(here("Data", "Processed", "rostralLmkSymm.rda"))
load(here("Data", "Processed", "neuroLmkSymm.rda"))
load(here("Data", "Processed", "fnpLmkSymm.rda"))
load(here("Data", "Processed", "mesLmkSymm.rda"))
load(here("Data", "Processed", "paLmkSymm.rda"))
load(file = here("Data", "Processed", "skullGPA.rda"))
load(file = here("Data", "Processed", "rostralGPA.rda"))
load(file = here("Data", "Processed", "neuroGPA.rda"))
load(file = here("Data", "Processed", "fnpGPA.rda"))
load(file = here("Data", "Processed", "mesGPA.rda"))
load(file = here("Data", "Processed", "paGPA.rda"))


# Load the labels produced in "GMM_GPA.Rmd"
pointLabels <- readRDS(here("Data", "Processed", "pointLabels.rds"))

#Load phylogeny - this is the composite tree from Rovinsky et al., 2021, with the fossil taxa added
# NB: Prothylacinus patagonicus is not included this tree
rawTree <- read.tree(here::here("Data", "phyloGMM.tre"))
```



#Allometry with Andrewsarchus is strong but related to Andrewsarchus being a high-leverage outlier.
```{r}
#shape

new.subset <-coords.subset(A = skullLmkSymm$symm.shape,
                           group = pointLabels$Taxon) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL

#reorder coords to fit the tree
meanCoords <- meanCoords[,,order(match(dimnames(meanCoords)[[3]],rawTree$tip.label))]


Csize <- as.numeric(skullGPA$Csize)
meanSkullCsize <- aggregate(Csize, 
                       by = list(pointLabels$Taxon), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL
rownames(meanSkullCsize) <- meanSkullCsize$Group.1
colnames(meanSkullCsize) <- c("Taxon", "Csize")
meanSkullCsize <- droplevels(meanSkullCsize)

#reorder Csize to fit the tree
meanSkullCsize <- meanSkullCsize[order(match(meanSkullCsize[,1],rawTree$tip.label)),]

#checks
rownames(meanSkullCsize) == rawTree$tip.label
dimnames(meanCoords)[[3]] == rawTree$tip.label
rownames(meanSkullCsize) == dimnames(meanCoords)[[3]]

allometry_gdf <- geomorph.data.frame(coords = meanCoords, Csize = meanSkullCsize$Csize, phy=rawTree)


allometry_fit <- procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdf)

summary(allometry_fit)

#Andrewsarchus is a high-leverage outlier
plot(allometry_fit)


```

#Allometry without Andrewsarchus - residuals are appropriately distributed
```{r}
#take out Andrewsarchus

rawTree <- drop.tip(phy = rawTree, tip = "Andrewsarchus_mongoliensis")

#shape

new.subset <-coords.subset(A = skullLmkSymm$symm.shape[,,-1],
                           group = droplevels( pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL

#reorder coords to fit the tree
meanCoords <- meanCoords[,,order(match(dimnames(meanCoords)[[3]],rawTree$tip.label))]



rawTree$tip.label == dimnames(meanCoords)[[3]]


#Csize without Andrewsarchus
Csize <- as.numeric(skullGPA$Csize)[-1]
meanSkullCsize <- aggregate(Csize, 
                       by = list(pointLabels$Taxon[-1]), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL

rownames(meanSkullCsize) <- meanSkullCsize$Group.1
colnames(meanSkullCsize) <- c("Taxon", "Csize")
meanSkullCsize <- droplevels(meanSkullCsize)

#reorder Csize to fit the tree
meanSkullCsize <- meanSkullCsize[order(match(meanSkullCsize[,1],rawTree$tip.label)),]

#checks
rownames(meanSkullCsize) == rawTree$tip.label
dimnames(meanCoords)[[3]] == rawTree$tip.label
rownames(meanSkullCsize) == dimnames(meanCoords)[[3]]


allometry_gdf <- geomorph.data.frame(coords = meanCoords, Csize = meanSkullCsize$Csize, phy=rawTree)


allometry_fit <- procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdf)

summary(allometry_fit)

#Here, the residuals are more appropriate
plot(allometry_fit)

Allom_PCA <- gm.prcomp(allometry_gdf$coords)


PCA_scores <-  as.data.frame(Allom_PCA$x)
Csize_taxa <- data.frame(Csize=meanSkullCsize$Csize, species = meanSkullCsize$Taxon)

#Check
Csize_taxa$species ==rownames(Allom_PCA$x)

Data_for_pgls <- cbind (PCA_scores, Csize_taxa)

Csize_PCA_frame <- comparative.data( rawTree, Data_for_pgls, names.col = species)

test <- pgls(Comp1~ log(Csize), data=Csize_PCA_frame)


summary(test)



```

#The first two PCAs are not associated with size, then it is smeared across PCs
```{r}
PCs_tested <- 10
pgls_PCA_table <- matrix(nrow=PCs_tested, ncol = 5)
colnames(pgls_PCA_table) <- c("Estimate",  "Std. Error",   "t value", "Rsq", "p")

for (i in 1:PCs_tested){
  
  temp <- summary(pgls( get(paste("Comp", i, sep="")) ~ log(Csize), data=Csize_PCA_frame))
  
 pgls_PCA_table[i,1:3]<- temp$coefficients[2,1:3]
 pgls_PCA_table[i,4]<- temp$r.squared
 pgls_PCA_table[i,5]<- temp$coefficients[2,4]

}

pgls_PCA_table[,c(1:3, 5)] <- round(pgls_PCA_table[,c(1:3, 5)], 3)
pgls_PCA_table[,4] <- round(pgls_PCA_table[,4], 2)

rownames(pgls_PCA_table)<- paste (rep("PC", PCs_tested), c(1:PCs_tested), sep="")

write.csv(pgls_PCA_table, file="./Figures/pgls_PCA_table.csv")


```


#We are justified in keeping Andrewsarchus in the PCA plots because it does not impact on the morphospaces much. The below recreates Fig. 1 without Andrewsarchus.


```{r Labels}
pointLabels$PtCol <- "#fff0e0"
pointLabels$PtCol[pointLabels$Clade == "Metatheria"] <- "#e0f0ff"
pointLabels$PtCol[pointLabels$Family == "Thylacinidae"] <- "#c87ef8" #purple
pointLabels$PtCol[pointLabels$Taxon == "Borhyaena_tuberata"] <- "#ca576c" #red - just to see position. 
pointLabels$PtCol[pointLabels$Family == "Canidae"] <- "#ffaa5c" #blue
pointLabels$PtCol[pointLabels$Taxon == "Canis_lupus"] <- "#233045" #navy - just to see position
pointLabels$PtCol[pointLabels$Taxon == "Andrewsarchus_mongoliensis"] <- "#72c496" #green - just to see position
pointLabels$Point <- 21
pointLabels$Point[pointLabels$Clade == "Metatheria"] <- 23
pointLabels$ID <- str_c(substr(pointLabels$Taxon,1,3), str_to_title(substr(sub(".*_", "", pointLabels$Taxon),1,3)) )
# For publication, we'll change those 'position' colours back to their clade colours and use text call-outs


# Gather mean value labels
meanLabels <- pointLabels
# This pulls out labels by species, and keeps the rest of the associated data columns
meanLabels <- meanLabels %>%
    distinct(Taxon, .keep_all = T)
row.names(meanLabels) <- meanLabels$Taxon
meanLabels <- meanLabels[order(meanLabels$Taxon, decreasing = F), ]
```

#### Skull without Andrewsarchus
```{r Skull Set}

new.subset <-coords.subset(A = skullLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



skullPCA <- gm.prcomp(meanCoords) 
skullPCPlot <- plot(skullPCA)
PCScores <- skullPCA$x[,1:10]
skullPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```

#### Rostrum
```{r Skull Set}
new.subset <-coords.subset(A = rostralLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



rostPCA <- gm.prcomp(meanCoords) 
rostPCPlot <- plot(rostPCA)
PCScores <- rostPCA$x[,1:10]
rostPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```
#### Neurocranium
```{r Skull Set}
new.subset <-coords.subset(A = neuroLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



neurPCA <- gm.prcomp(meanCoords) 
neurPCPlot <- plot(neurPCA)
PCScores <- neurPCA$x[,1:10]
neurPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```

#Note that PC2 is flipped but otherwise very similar to Fig. 1
```{r Plot}

pdf(file="./Figures/Supp1_AMong_removed.pdf", width=10, height=10)

par(mfrow=c(2,2))

plot(NULL,
     xlim = range(skullPlotData$Comp1),
     ylim = -range(skullPlotData$Comp2),
     lwd = 0.5,
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Full Cranium")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(skullPlotData$Comp2 ~ skullPlotData$Comp1, 
     pch = skullPlotData$Point, 
     col = "#000000", 
     bg = skullPlotData$PtCol, 
     cex = 2)


#rostrum

plot(NULL,
     xlim = range(rostPlotData$Comp1),
     ylim = -range(rostPlotData$Comp2),
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Rostrum")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(rostPlotData$Comp2 ~ rostPlotData$Comp1, 
     pch = rostPlotData$Point, 
     col = "#000000", 
     bg = rostPlotData$PtCol, 
     cex = 2)



#Neurocranium

plot(NULL,
     xlim = range(neurPlotData$Comp1),
     ylim = -range(neurPlotData$Comp2),
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Neurocranium")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(neurPlotData$Comp2 ~ neurPlotData$Comp1, 
     pch = neurPlotData$Point, 
     col = "#000000", 
     bg = neurPlotData$PtCol, 
     cex = 2)

legend("bottomleft",
       legend = expression("Placentals", "Canids",
                  "Metatherians", "Thylacinidae", italic("Borhyaena"), italic("C. lupus")),
       pch = c(21, 21, 23, 23),
       col = c("#000000", "#000000", "#000000", "#000000"),
       pt.bg = c("#fff0e0", "#ffaa5c", "#e0f0ff", "#c87ef8", "#ca576c", "black"),
       bty = "n",
       pt.cex = 2,
       cex = 1,
       text.col = "black",
       horiz = F ,
       inset = c(0.00, 0.0))

plot.new()



dev.off()

```


#Allometry of the various partitions - all without _Andrewsarchus_

```{r}

#take out Andrewsarchus

rawTree <- drop.tip(phy = rawTree, tip = "Andrewsarchus_mongoliensis")

#take out Andrewsarchus for all partitions

partition_list_raw <- list(skull=skullLmkSymm, rostral=rostralLmkSymm, neuro=neuroLmkSymm, fnp=fnpLmkSymm, pa=paLmkSymm, mes=mesLmkSymm)
partition_list_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

meancoords_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for(i in 1:length(partition_list_noAnd)){

partition_list_noAnd[[i]] <-coords.subset(A = partition_list_raw[[i]]$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meancoords_noAnd[[i]] <- sapply(partition_list_noAnd[[i]], 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL
#reorder coords to fit the tree
meancoords_noAnd[[i]] <- meancoords_noAnd[[i]][,,order(match(dimnames(meancoords_noAnd[[i]])[[3]],rawTree$tip.label))]

}

#Check
rawTree$tip.label == dimnames(meancoords_noAnd[[3]]) [[3]]

#Csizes

csizes_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)
meanSkullCsize_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for(i in 1:length(partition_list_noAnd)){
  
csizes_noAnd[[i]] <- as.numeric(get(paste(names(csizes_noAnd)[i], "GPA", sep=""))$Csize)[-1]  
 
meanSkullCsize_noAnd[[i]] <- aggregate(csizes_noAnd[[i]], 
                       by = list(pointLabels$Taxon[-1]), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL

rownames(meanSkullCsize_noAnd[[i]]) <- meanSkullCsize$Group.1
colnames(meanSkullCsize_noAnd[[i]]) <- c("Taxon", "Csize")
meanSkullCsize_noAnd[[i]] <- droplevels(meanSkullCsize_noAnd[[i]]) 
 
#rearrange to fit tree
meanSkullCsize_noAnd[[i]] <- meanSkullCsize_noAnd[[i]][order(match(meanSkullCsize_noAnd[[i]][,1],rawTree$tip.label)),] 
  
  
  
}

#Now for the allometry analyses

#quick check that all lines up
meanSkullCsize_noAnd[[1]]$Taxon==dimnames(meancoords_noAnd$skull)[[3]]
dimnames(meancoords_noAnd$skull)[[3]] == rawTree$tip.label

allometry_gdfs <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)
allometry_fits <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for (i in 1:length(partition_list_noAnd)){
  
allometry_gdfs[[i]] <- geomorph.data.frame(coords = meancoords_noAnd[[i]], Csize = meanSkullCsize_noAnd[[i]]$Csize, taxon = meanSkullCsize_noAnd[[i]]$Taxon, phy=rawTree)  

allometry_fits[[i]] <- summary(procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdfs[[i]]))

}

#make output table

allometry_output <- matrix(nrow=length(partition_list_noAnd), ncol=length(allometry_fits$skull$table[1,]))
rownames(allometry_output) <- names(partition_list_noAnd)
colnames(allometry_output) <- colnames(allometry_fits$skull$table[1,] )

for (i in 1:length(partition_list_noAnd)){
  
 allometry_output[i,]<- unlist(allometry_fits[[i]]$table[1,])
  
  
}

#just so the processed talbe doesn't overwrite the previous
allometry_output_table <- allometry_output

allometry_output_table[,c(2,3,5,6,7)] <- sprintf("%.3f", allometry_output[,c(2,3,5,6,7)] )
allometry_output_table[,c(4)] <- sprintf("%.2f", allometry_output[,c(4)])

allometry_output_table<- allometry_output_table[,-c(1,2)]

allometry_output_table 

?procD.pgls

write.csv(allometry_output_table, file = "./Figures/Allometry_output.csv")

```

#adding the PCA thingy

```{r}

Allom_PCAs <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)
 temp <-NULL

for (i in 1:length(Allom_PCAs)){
  
temp<- gm.prcomp(allometry_gdfs[[i]]$coords)
Allom_PCAs[[i]]$PCA_Csize <- cbind(as.data.frame(temp$x), as.numeric(allometry_gdfs[[i]]$Csize), as.character(allometry_gdfs[[i]]$taxon))

#soft coding this, messy but safer in case the pc numbers change
colnames(Allom_PCAs[[i]]$PCA_Csize)[length(colnames(Allom_PCAs[[i]]$PCA_Csize))-1] <- "Csize"
colnames(Allom_PCAs[[i]]$PCA_Csize)[length(colnames(Allom_PCAs[[i]]$PCA_Csize))]<- "species"
Allom_PCAs[[i]]$phy <-allometry_gdfs[[i]]$phy 


  
}



#Check checking
rownames(Allom_PCAs$rostral$PCA_Csize) == Allom_PCAs$neuro$phy$tip.label
Allom_PCAs$rostral$PCA_Csize$species == Allom_PCAs$neuro$phy$tip.label
  

Csize_PCA_frames <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for(i in 1:length(Csize_PCA_frames)){

Csize_PCA_frames [[i]] <- comparative.data( rawTree , 
                                            Allom_PCAs[[i]]$PCA_Csize, 
                                            'species')

}


# #This is a test to make sure this really works - the result is the same as a previous analysis that worked


# test_data_frame <- Allom_PCAs$skull$PCA_Csize[,1:(length(colnames(Allom_PCAs[[i]]$PCA_Csize)))]
# test_tree <- rawTree
# testnames <- as.character(Allom_PCAs[[i]]$PCA_Csize[length(colnames(Allom_PCAs[[i]]$PCA_Csize))])
# 
# test_comparative_data <- comparative.data(rawTree, 
#                                            Allom_PCAs$skull$PCA_Csize[,1:(length(colnames(Allom_PCAs[[i]]$PCA_Csize)))], 
#                                            'species' )
# 
# 
# 

# test <- pgls(Comp1 ~ log(Csize), data =Csize_PCA_frames$skull )
# summary(test)

#Setting up output tables
PCs_tested <- 10
pgls_PCA_table <- matrix(nrow=PCs_tested, ncol = 5)
colnames(pgls_PCA_table) <- c("Estimate",  "Std. Error",   "t value", "Rsq", "p")
rownames(pgls_PCA_table)<- paste (rep("PC", PCs_tested), c(1:PCs_tested), sep="")


comparative_data_PCA_outputs <- list(skull=pgls_PCA_table, rostral=pgls_PCA_table, neuro=pgls_PCA_table, fnp=pgls_PCA_table, pa=pgls_PCA_table, mes=pgls_PCA_table)


for (i in 1: length(Csize_PCA_frames)){
  
  for (j in 1: PCs_tested){
  
    
  temp <- summary(pgls( get(paste("Comp", j, sep="")) ~ log(Csize), data=Csize_PCA_frames[[i]]))
  
  comparative_data_PCA_outputs[[i]] [j,1:3]<- sprintf("%.3f",temp$coefficients[2,1:3])
 comparative_data_PCA_outputs[[i]][j,4]<- sprintf("%.2f", temp$r.squared)
 comparative_data_PCA_outputs[[i]][j,5]<- sprintf("%.3f", temp$coefficients[2,4])
 
  }
  
}

  

comparative_data_PCA_outputs$neuro

#OK packaging all of this into one huge table. stringsAsFactors = FALSE makes it so not everything turns into a character

PCA_All_outputs <-data.frame( rbind(c("Cranium", rep(NA, ncol(comparative_data_PCA_outputs$skull)-2)), comparative_data_PCA_outputs$skull[,-1],
      c("Rostrum", rep(NA, ncol(comparative_data_PCA_outputs$rostral)-2)), comparative_data_PCA_outputs$rostral[,-1],
      c("Neurocranium", rep(NA, ncol(comparative_data_PCA_outputs$neuro)-2)), comparative_data_PCA_outputs$neuro[,-1],
      c("Frontonasal Process", rep(NA, ncol(comparative_data_PCA_outputs$fnp)-2)), comparative_data_PCA_outputs$fnp[,-1],
      c("First Pharyngeal Arch", rep(NA, ncol(comparative_data_PCA_outputs$pa)-2)), comparative_data_PCA_outputs$pa[,-1],
      c("Paraxial Mesoderm", rep(NA, ncol(comparative_data_PCA_outputs$mes)-2)), comparative_data_PCA_outputs$mes[,-1]
),stringsAsFactors = FALSE)



write.csv(PCA_All_outputs, file="./Figures/pgls_PCA_tables.csv")



#because this is a lot of coding and potential for human error, I'm here running the PCAs work - it checks out.

# plot(Csize_PCA_frames$skull$data$Comp2~Csize_PCA_frames$skull$data$Comp1, asp=TRUE)
# text(Csize_PCA_frames$skull$data$Comp2~Csize_PCA_frames$skull$data$Comp1, asp=TRUE, labels=rownames(Csize_PCA_frames$skull$data) )
# 

```


# ```{r}
# 
# #this is the code for just one variable
# 
# 
# PCs_tested <- 10
# pgls_PCA_table <- matrix(nrow=PCs_tested, ncol = 5)
# colnames(pgls_PCA_table) <- c("Estimate",  "Std. Error",   "t value", "Rsq", "p")
# 
# for (i in 1:PCs_tested){
# 
#   temp <- summary(pgls( get(paste("Comp", i, sep="")) ~ log(Csize), data=Csize_PCA_frame))
# 
#  pgls_PCA_table[i,1:3]<- temp$coefficients[2,1:3]
#  pgls_PCA_table[i,4]<- temp$r.squared
#  pgls_PCA_table[i,5]<- temp$coefficients[2,4]
# 
# }
# 
# pgls_PCA_table[,c(1:3, 5)] <- round(pgls_PCA_table[,c(1:3, 5)], 3)
# pgls_PCA_table[,4] <- round(pgls_PCA_table[,4], 2)
# 
# rownames(pgls_PCA_table)<- paste (rep("PC", PCs_tested), c(1:PCs_tested), sep="")
# 
# write.csv(pgls_PCA_table, file="./Figures/pgls_PCA_table.csv")
# 
# 
# ```



#Making Table 1 - allometry
```{r}

allometry_output_table

PC12_list <- matrix(nrow=length(comparative_data_PCA_outputs), ncol=4)

for (i in 1:length(comparative_data_PCA_outputs)){
  
PC12_list[i,] <- c (comparative_data_PCA_outputs[[i]][c(1), c(4,5)], comparative_data_PCA_outputs[[i]][c(2), c(4,5)])
}


Table_1 <- cbind(allometry_output_table, PC12_list)

colnames(Table_1)[c(5:9)]<- c("p", "PC1_Rsq", "PC1_p", "PC2_Rsq", "PC2_p")


Table_1 <- data.frame(Table_1, stringsAsFactors = FALSE)

write.csv(Table_1, file="./Figures/Table_1_Allometry.csv")

```





