---
title: "01_01_Allometry_check"
author: "Vera Weisbecker"
date: "2024-05-03"
output: html_document
editor_options:
  chunk_output_type: console
---


```{r}

library(geomorph)  # CRAN v4.0.10
library(ape)       # CRAN v5.8-1       
library(caper)     # CRAN v1.0.3    
library(tidyverse) # CRAN v2.0.0 
library(svglite)   # CRAN v2.2.1 
library(here)      # CRAN v1.0.1     

load(here("Data", "Processed", "skullLmkSymm.rda"))
load(here("Data", "Processed", "rostralLmkSymm.rda"))
load(here("Data", "Processed", "neuroLmkSymm.rda"))
load(here("Data", "Processed", "fnpLmkSymm.rda"))
load(here("Data", "Processed", "mesLmkSymm.rda"))
load(here("Data", "Processed", "paLmkSymm.rda"))
load(file = here("Data", "Processed", "skullGPA.rda"))
load(file = here("Data", "Processed", "rostralGPA.rda"))
load(file = here("Data", "Processed", "neuroGPA.rda"))
load(file = here("Data", "Processed", "fnpGPA.rda"))
load(file = here("Data", "Processed", "mesGPA.rda"))
load(file = here("Data", "Processed", "paGPA.rda"))


# Load the labels produced in "GMM_GPA.Rmd"
pointLabels <- readRDS(here("Data", "Processed", "pointLabels.rds"))

#Load phylogeny - this is the composite tree from Rovinsky et al., 2021, with the fossil taxa added
# NB: Prothylacinus patagonicus is not included this tree
rawTree <- read.tree(here::here("Data", "phyloGMM.tre"))
```



#Allometry with Andrewsarchus is strong but related to Andrewsarchus being a high-leverage outlier.
```{r}
#shape

new.subset <-coords.subset(A = skullLmkSymm$symm.shape,
                           group = pointLabels$Taxon) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL

#reorder coords to fit the tree
meanCoords <- meanCoords[,,order(match(dimnames(meanCoords)[[3]],rawTree$tip.label))]


Csize <- as.numeric(skullGPA$Csize)
meanSkullCsize <- aggregate(Csize, 
                       by = list(pointLabels$Taxon), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL
rownames(meanSkullCsize) <- meanSkullCsize$Group.1
colnames(meanSkullCsize) <- c("Taxon", "Csize")
meanSkullCsize <- droplevels(meanSkullCsize)

#reorder Csize to fit the tree
meanSkullCsize <- meanSkullCsize[order(match(meanSkullCsize[,1],rawTree$tip.label)),]

#checks
rownames(meanSkullCsize) == rawTree$tip.label
dimnames(meanCoords)[[3]] == rawTree$tip.label
rownames(meanSkullCsize) == dimnames(meanCoords)[[3]]

allometry_gdf <- geomorph.data.frame(coords = meanCoords, Csize = meanSkullCsize$Csize, phy=rawTree)


allometry_fit <- procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdf)

par(mfrow=c(2,2))

summary(allometry_fit)

#Andrewsarchus is a high-leverage outlier
plot(allometry_fit)


```

#Allometry without Andrewsarchus - residuals are appropriately distributed
```{r}
#take out Andrewsarchus

rawTree <- drop.tip(phy = rawTree, tip = "Andrewsarchus_mongoliensis")

#shape

new.subset <-coords.subset(A = skullLmkSymm$symm.shape[,,-1],
                           group = droplevels( pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL

#reorder coords to fit the tree
meanCoords <- meanCoords[,,order(match(dimnames(meanCoords)[[3]],rawTree$tip.label))]



rawTree$tip.label == dimnames(meanCoords)[[3]]


#Csize without Andrewsarchus
Csize <- as.numeric(skullGPA$Csize)[-1]
meanSkullCsize <- aggregate(Csize, 
                       by = list(pointLabels$Taxon[-1]), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL

rownames(meanSkullCsize) <- meanSkullCsize$Group.1
colnames(meanSkullCsize) <- c("Taxon", "Csize")
meanSkullCsize <- droplevels(meanSkullCsize)

#reorder Csize to fit the tree
meanSkullCsize <- meanSkullCsize[order(match(meanSkullCsize[,1],rawTree$tip.label)),]

#checks
rownames(meanSkullCsize) == rawTree$tip.label
dimnames(meanCoords)[[3]] == rawTree$tip.label
rownames(meanSkullCsize) == dimnames(meanCoords)[[3]]


allometry_gdf <- geomorph.data.frame(coords = meanCoords, Csize = meanSkullCsize$Csize, phy=rawTree)


allometry_fit <- procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdf)

summary(allometry_fit)

#Here, the residuals are more appropriate
plot(allometry_fit)

Allom_PCA <- gm.prcomp(allometry_gdf$coords)


PCA_scores <-  as.data.frame(Allom_PCA$x)
Csize_taxa <- data.frame(Csize=meanSkullCsize$Csize, species = meanSkullCsize$Taxon)

#Check
Csize_taxa$species ==rownames(Allom_PCA$x)

Data_for_pgls <- cbind (PCA_scores, Csize_taxa)

Csize_PCA_frame <- comparative.data( rawTree, Data_for_pgls, names.col = species)

test <- pgls(Comp1~ log(Csize), data=Csize_PCA_frame)


summary(test)



```


#We are justified in keeping Andrewsarchus in the PCA plots because it does not impact on the morphospaces much. The below recreates Fig. 1 without Andrewsarchus.


```{r Labels}
pointLabels$PtCol <- "#fff0e0"
pointLabels$PtCol[pointLabels$Clade == "Metatheria"] <- "#e0f0ff"
pointLabels$PtCol[pointLabels$Family == "Thylacinidae"] <- "#c87ef8" #purple
pointLabels$PtCol[pointLabels$Taxon == "Borhyaena_tuberata"] <- "#ca576c" #red - just to see position. 
pointLabels$PtCol[pointLabels$Family == "Canidae"] <- "#ffaa5c" #blue
pointLabels$PtCol[pointLabels$Taxon == "Canis_lupus"] <- "#233045" #navy - just to see position
pointLabels$PtCol[pointLabels$Taxon == "Andrewsarchus_mongoliensis"] <- "#72c496" #green - just to see position
pointLabels$Point <- 21
pointLabels$Point[pointLabels$Clade == "Metatheria"] <- 23
pointLabels$ID <- str_c(substr(pointLabels$Taxon,1,3), str_to_title(substr(sub(".*_", "", pointLabels$Taxon),1,3)) )
# For publication, we'll change those 'position' colours back to their clade colours and use text call-outs


# Gather mean value labels
meanLabels <- pointLabels
# This pulls out labels by species, and keeps the rest of the associated data columns
meanLabels <- meanLabels %>%
    distinct(Taxon, .keep_all = T)
row.names(meanLabels) <- meanLabels$Taxon
meanLabels <- meanLabels[order(meanLabels$Taxon, decreasing = F), ]
```

#### Skull without Andrewsarchus
```{r Skull Set}

new.subset <-coords.subset(A = skullLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



skullPCA <- gm.prcomp(meanCoords) 
skullPCPlot <- plot(skullPCA)
PCScores <- skullPCA$x[,1:10]
skullPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```

#### Rostrum without Andrewsarchus
```{r Skull Set}
new.subset <-coords.subset(A = rostralLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



rostPCA <- gm.prcomp(meanCoords) 
rostPCPlot <- plot(rostPCA)
PCScores <- rostPCA$x[,1:10]
rostPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```
#### Neurocranium without Andrewsarchus
```{r Skull Set}
new.subset <-coords.subset(A = neuroLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



neurPCA <- gm.prcomp(meanCoords) 
neurPCPlot <- plot(neurPCA)
PCScores <- neurPCA$x[,1:10]
neurPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```

#Note that PC2 is flipped but otherwise very similar to Fig. 1
```{r Plot}

pdf(file="./Figures/Supp1_AMong_removed.pdf", width=10, height=10)

par(mfrow=c(2,2))

plot(NULL,
     xlim = range(skullPlotData$Comp1),
     ylim = -range(skullPlotData$Comp2),
     lwd = 0.5,
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Full Cranium")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(skullPlotData$Comp2 ~ skullPlotData$Comp1, 
     pch = skullPlotData$Point, 
     col = "#000000", 
     bg = skullPlotData$PtCol, 
     cex = 2)


#rostrum

plot(NULL,
     xlim = range(rostPlotData$Comp1),
     ylim = -range(rostPlotData$Comp2),
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Rostrum")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(rostPlotData$Comp2 ~ rostPlotData$Comp1, 
     pch = rostPlotData$Point, 
     col = "#000000", 
     bg = rostPlotData$PtCol, 
     cex = 2)



#Neurocranium

plot(NULL,
     xlim = range(neurPlotData$Comp1),
     ylim = -range(neurPlotData$Comp2),
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Neurocranium")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(neurPlotData$Comp2 ~ neurPlotData$Comp1, 
     pch = neurPlotData$Point, 
     col = "#000000", 
     bg = neurPlotData$PtCol, 
     cex = 2)

legend("bottomleft",
       legend = expression("Placentals", "Canids",
                  "Metatherians", "Thylacinidae", italic("Borhyaena"), italic("C. lupus")),
       pch = c(21, 21, 23, 23),
       col = c("#000000", "#000000", "#000000", "#000000"),
       pt.bg = c("#fff0e0", "#ffaa5c", "#e0f0ff", "#c87ef8", "#ca576c", "black"),
       bty = "n",
       pt.cex = 2,
       cex = 1,
       text.col = "black",
       horiz = F ,
       inset = c(0.00, 0.0))

plot.new()



dev.off()

```


#Allometry of the various partitions - all without Andrewsarchus

```{r}

#take out Andrewsarchus

rawTree <- drop.tip(phy = rawTree, tip = "Andrewsarchus_mongoliensis")

#take out Andrewsarchus for all partitions

partition_list_raw <- list(skull=skullLmkSymm, rostral=rostralLmkSymm, neuro=neuroLmkSymm, fnp=fnpLmkSymm, pa=paLmkSymm, mes=mesLmkSymm)
partition_list_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

meancoords_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for(i in 1:length(partition_list_noAnd)){

partition_list_noAnd[[i]] <-coords.subset(A = partition_list_raw[[i]]$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meancoords_noAnd[[i]] <- sapply(partition_list_noAnd[[i]], 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL
#reorder coords to fit the tree
meancoords_noAnd[[i]] <- meancoords_noAnd[[i]][,,order(match(dimnames(meancoords_noAnd[[i]])[[3]],rawTree$tip.label))]

}

#Check
rawTree$tip.label == dimnames(meancoords_noAnd[[3]]) [[3]]

#Csizes

csizes_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)
meanSkullCsize_noAnd <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for(i in 1:length(partition_list_noAnd)){
  
csizes_noAnd[[i]] <- as.numeric(get(paste(names(csizes_noAnd)[i], "GPA", sep=""))$Csize)[-1]  
 
meanSkullCsize_noAnd[[i]] <- aggregate(csizes_noAnd[[i]], 
                       by = list(pointLabels$Taxon[-1]), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL

rownames(meanSkullCsize_noAnd[[i]]) <- meanSkullCsize$Group.1
colnames(meanSkullCsize_noAnd[[i]]) <- c("Taxon", "Csize")
meanSkullCsize_noAnd[[i]] <- droplevels(meanSkullCsize_noAnd[[i]]) 
 
#rearrange to fit tree
meanSkullCsize_noAnd[[i]] <- meanSkullCsize_noAnd[[i]][order(match(meanSkullCsize_noAnd[[i]][,1],rawTree$tip.label)),] 
  
  
  
}

#Now for the allometry analyses

#quick check that all lines up
meanSkullCsize_noAnd[[1]]$Taxon==dimnames(meancoords_noAnd$skull)[[3]]
dimnames(meancoords_noAnd$skull)[[3]] == rawTree$tip.label

allometry_gdfs <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)
allometry_fits <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for (i in 1:length(partition_list_noAnd)){
  
allometry_gdfs[[i]] <- geomorph.data.frame(coords = meancoords_noAnd[[i]], Csize = meanSkullCsize_noAnd[[i]]$Csize, taxon = meanSkullCsize_noAnd[[i]]$Taxon, phy=rawTree)  

allometry_fits[[i]] <- summary(procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdfs[[i]]))

}

#make output table

allometry_output <- matrix(nrow=length(partition_list_noAnd), ncol=length(allometry_fits$skull$table[1,]))
rownames(allometry_output) <- names(partition_list_noAnd)
colnames(allometry_output) <- colnames(allometry_fits$skull$table[1,] )

for (i in 1:length(partition_list_noAnd)){
  
 allometry_output[i,]<- unlist(allometry_fits[[i]]$table[1,])
  
  
}

#just so the processed talbe doesn't overwrite the previous
allometry_output_table <- allometry_output

allometry_output_table[,c(2,3,5,6,7)] <- sprintf("%.3f", allometry_output[,c(2,3,5,6,7)] )
allometry_output_table[,c(4)] <- sprintf("%.2f", allometry_output[,c(4)])

allometry_output_table<- allometry_output_table[,-c(1,2)]

allometry_output_table 


write.csv(allometry_output_table, file = "./Figures/Allometry_output.csv")

```



#adding the PCA 

```{r}

Allom_PCAs <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)
 temp <-NULL

for (i in 1:length(Allom_PCAs)){
  
temp<- gm.prcomp(allometry_gdfs[[i]]$coords)
Allom_PCAs[[i]]$PCA_Csize <- cbind(as.data.frame(temp$x), as.numeric(allometry_gdfs[[i]]$Csize), as.character(allometry_gdfs[[i]]$taxon))

#soft coding this, messy but safer in case the pc numbers change
colnames(Allom_PCAs[[i]]$PCA_Csize)[length(colnames(Allom_PCAs[[i]]$PCA_Csize))-1] <- "Csize"
colnames(Allom_PCAs[[i]]$PCA_Csize)[length(colnames(Allom_PCAs[[i]]$PCA_Csize))]<- "species"
Allom_PCAs[[i]]$phy <-allometry_gdfs[[i]]$phy 


  
}



#Check checking
rownames(Allom_PCAs$rostral$PCA_Csize) == Allom_PCAs$neuro$phy$tip.label
Allom_PCAs$rostral$PCA_Csize$species == Allom_PCAs$neuro$phy$tip.label
  

Csize_PCA_frames <- list(skull=NULL, rostral=NULL, neuro=NULL, fnp=NULL, pa=NULL, mes=NULL)

for(i in 1:length(Csize_PCA_frames)){

Csize_PCA_frames [[i]] <- comparative.data( rawTree , 
                                            Allom_PCAs[[i]]$PCA_Csize, 
                                            'species')

}


#Setting up output tables
PCs_tested <- 10
pgls_PCA_table <- matrix(nrow=PCs_tested, ncol = 5)
colnames(pgls_PCA_table) <- c("Estimate",  "Std. Error",   "t value", "Rsq", "p")
rownames(pgls_PCA_table)<- paste (rep("PC", PCs_tested), c(1:PCs_tested), sep="")


comparative_data_PCA_outputs <- list(skull=pgls_PCA_table, rostral=pgls_PCA_table, neuro=pgls_PCA_table, fnp=pgls_PCA_table, pa=pgls_PCA_table, mes=pgls_PCA_table)


for (i in 1: length(Csize_PCA_frames)){
  
  for (j in 1: PCs_tested){
  
    
  temp <- summary(pgls( get(paste("Comp", j, sep="")) ~ log(Csize), data=Csize_PCA_frames[[i]]))
  
  comparative_data_PCA_outputs[[i]] [j,1:3]<- sprintf("%.3f",temp$coefficients[2,1:3])
 comparative_data_PCA_outputs[[i]][j,4]<- sprintf("%.2f", temp$r.squared)
 comparative_data_PCA_outputs[[i]][j,5]<- sprintf("%.3f", temp$coefficients[2,4])
 
  }
  
}

  

comparative_data_PCA_outputs$neurocranium

#OK packaging all of this into one huge table. stringsAsFactors = FALSE makes it so not everything turns into a character

PCA_All_outputs <-data.frame( rbind(c("Cranium", rep(NA, ncol(comparative_data_PCA_outputs$skull)-2)), comparative_data_PCA_outputs$skull[,-1],
      c("Rostrum", rep(NA, ncol(comparative_data_PCA_outputs$rostral)-2)), comparative_data_PCA_outputs$rostral[,-1],
      c("Neurocranium", rep(NA, ncol(comparative_data_PCA_outputs$neuro)-2)), comparative_data_PCA_outputs$neuro[,-1],
      c("Frontonasal Process", rep(NA, ncol(comparative_data_PCA_outputs$fnp)-2)), comparative_data_PCA_outputs$fnp[,-1],
      c("First Pharyngeal Arch", rep(NA, ncol(comparative_data_PCA_outputs$pa)-2)), comparative_data_PCA_outputs$pa[,-1],
      c("Paraxial Mesoderm", rep(NA, ncol(comparative_data_PCA_outputs$mes)-2)), comparative_data_PCA_outputs$mes[,-1]
),stringsAsFactors = FALSE)



write.csv(PCA_All_outputs, file="./Figures/pgls_PCA_tables.csv")



#because this is a lot of coding and potential for human error, I'm here running the PCAs work - it checks out.

# plot(Csize_PCA_frames$skull$data$Comp2~Csize_PCA_frames$skull$data$Comp1, asp=TRUE)
# text(Csize_PCA_frames$skull$data$Comp2~Csize_PCA_frames$skull$data$Comp1, asp=TRUE, labels=rownames(Csize_PCA_frames$skull$data) )
# 

```


#Making Table 1 - allometry
```{r}

allometry_output_table

PC12_list <- matrix(nrow=length(comparative_data_PCA_outputs), ncol=4)

for (i in 1:length(comparative_data_PCA_outputs)){
  
PC12_list[i,] <- c (comparative_data_PCA_outputs[[i]][c(1), c(4,5)], comparative_data_PCA_outputs[[i]][c(2), c(4,5)])
}


Table_1 <- cbind(allometry_output_table, PC12_list)

colnames(Table_1)[c(5:9)]<- c("p", "PC1_Rsq", "PC1_p", "PC2_Rsq", "PC2_p")


Table_1 <- data.frame(Table_1, stringsAsFactors = FALSE)

write.csv(Table_1, file="./Figures/Table_1_Allometry.csv")

```



#Allometry of separate clades,The most appropriate for this dataset are marsupials, feliforms, and canids. W

#create allometry ANCOVAs with clades as factor. 

```{r}

#We already have all partitions - these are allometry_gdfs. Now we have to split these into allometry_gdfs for each clade. Starting off with just one partition

#taxa match when removing Andrewsarchus from mean labels, but need to be reordered. Also Andrewsarchus needs removing as a level after removal.
meanLabels_noAnd <-meanLabels[-2,]
meanLabels_noAnd$Taxon <- droplevels (meanLabels_noAnd$Taxon)
allometry_gdfs$skull$taxon %in% meanLabels_noAnd[-2,]$Taxon
allometry_gdfs$skull$taxon == meanLabels_noAnd$Taxon

#So we now reorder the labels according to the gdf labels
Labels_allometry_partitions <- meanLabels_noAnd[match(allometry_gdfs$skull$taxon,meanLabels_noAnd$Taxon),]

#Now the labels are the right order and gdf taxon order remains unchanged and the trees in the gdfs also talks to the labels
allometry_gdfs$skull$taxon==Labels_allometry_partitions$Taxon

allometry_gdfs$skull$phy$tip.label == Labels_allometry_partitions$Taxon

#add clade information to gdfs and drop whippomorphan (Artiodactyla) level
allometry_clade_gdfs <- allometry_gdfs
allometry_clade_fits <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, Fnp=NULL, Pa=NULL, Mes=NULL)

for (i in 1:length( allometry_gdfs)){

  
  allometry_clade_gdfs[[i]]$labels <- Labels_allometry_partitions
  allometry_clade_gdfs[[i]]$labels$Clade <- droplevels(allometry_clade_gdfs[[i]]$labels$Clade)
  names(allometry_clade_gdfs[[i]]$Csize) <- dimnames(allometry_clade_gdfs[[i]]$coords)[[3]]
}

#all 58, as it should be, and all data contained in here
allometry_clade_gdfs[[i]]$phy
dimnames(allometry_clade_gdfs[[i]]$coords)[[3]]
allometry_clade_gdfs[[i]]$Csize
nrow(allometry_clade_gdfs[[i]]$labels)
names(allometry_clade_gdfs)

allometry_clade_gdfs$skull$labels



rawtree_clades <- rawTree

#If you want, here you can take borhyaena from the tree, make sure it talks to the gdf etc
#drop.tip (rawTree, which(as.vector(Labels_allometry_partitions$Family) == "Borhyaenidae")) 
#rawtree_clades$tip.label == dimnames(allometry_clade_gdfs[[1]]$coords)[[3]]
#plot(rawtree_clades)

#Tests below- always check for small mistakes that throw everything. For reasons I do NOT understand (grr!) I can't use the data= at the end and have to spell out the whole address. Doesn't really matter but I could swear that this didn't use to work and it had to be the other way!

#testfit <- summary(procD.pgls(allometry_clade_gdfs$skull$coords ~ log(allometry_clade_gdfs$skull$Csize), phy=rawtree_clades))

#just checking that the thylacine is still one of the largest-skulled species in the sample: 

#allometry_clade_gdfs$skull$Csize[order(log(allometry_clade_gdfs$skull$Csize))]

#also making sure the shape still is associated with the right species - all good, hashing out
#spheres3d( allometry_clade_gdfs$skull$coords[,,which (dimnames(allometry_clade_gdfs$skull$coords)[[3]]=="Thylacinus_cynocephalus")], radius = 0.001, col="hotpink")

#spheres3d( allometry_clade_gdfs$rostral$coords[,,which (dimnames(allometry_clade_gdfs$rostral$coords)[[3]]=="Thylacinus_cynocephalus")], radius = 0.001, col="hotpink")

```

# moving on to by-clade analysis, using LM because pgls would remove this signal

```{r}

#Caniforms, feliforms, marsupials;  labels$Clade . Note that we had previously run this with Borhyaena removed (it is a non-marsupial so we were on the fence about including it), but the results are nearly the same with slightly smaller (0.01 max different Rsquareds)


for (i in 1: length(allometry_clade_gdfs)){
  
  allometry_clade_fits[[i]] <- summary(procD.lm(allometry_clade_gdfs[[i]]$coords ~ log(allometry_clade_gdfs[[i]]$Csize)* allometry_clade_gdfs[[i]]$labels$Clade))
      
}


#Matrix with four rows per comparison (partition name, size, clade and interaction)
Allometry_table_clade <- matrix (nrow=(length(allometry_clade_fits)*4), ncol = 5)

#the below code needs this line to start
rownames(Allometry_table_clade)<- rep (names(allometry_clade_fits), each=4)

for (i in 2: length(allometry_clade_fits)){
#need to start off with one  
rownames(Allometry_table_clade)[1:4] <- c(names(allometry_clade_fits)[1], "log Csize", "Clade", "Clade:Csize")
#then add to the previous - there must be a nicer way of coding this, but I haven't found it and this is still better than repeated pastes
rownames(Allometry_table_clade)[(((i-1)*4)+1):(i*4)]<- c(names(allometry_clade_fits)[i], "log Csize", "Clade", "Clade:Csize")

}

colnames(Allometry_table_clade)<-c("SS", "RSq", "F", "Z", "p" )

#Now fill in the results

for (i in 2: length (allometry_clade_fits)){
  
#again, 1 needs separate treatment
  #SS
  Allometry_table_clade[2:4,1] <- round(allometry_clade_fits[[1]]$table$SS[1:3], 3)
  #Rsq
  Allometry_table_clade[2:4,2] <- round(allometry_clade_fits[[1]]$table$Rsq[1:3], 2)
  #F
  Allometry_table_clade[2:4,3] <- round(allometry_clade_fits[[1]]$table$F[1:3], 3)
  #F
  Allometry_table_clade[2:4,4] <- round(allometry_clade_fits[[1]]$table$Z[1:3], 3)
  #p
  Allometry_table_clade[2:4,5] <- round(allometry_clade_fits[[1]]$table$`Pr(>F)`[1:3], 3)
  
  #now all the others. i-1*4+2 here (not +1) because the row with the name of teh partition is blank
  #SS
  Allometry_table_clade[(((i-1)*4)+2):(i*4),1] <- round(allometry_clade_fits[[i]]$table$SS[1:3], 3)
  #Rsq
  Allometry_table_clade[(((i-1)*4)+2):(i*4),2] <- round(allometry_clade_fits[[i]]$table$Rsq[1:3], 2)
  #F
  Allometry_table_clade[(((i-1)*4)+2):(i*4),3] <- round(allometry_clade_fits[[i]]$table$F[1:3], 3)
  #F
  Allometry_table_clade[(((i-1)*4)+2):(i*4),4] <- round(allometry_clade_fits[[i]]$table$Z[1:3], 3)
  #p
  Allometry_table_clade[(((i-1)*4)+2):(i*4),5] <- round(allometry_clade_fits[[i]]$table$`Pr(>F)`[1:3], 3)
  
  }


write.csv(Allometry_table_clade, file="./Figures/Allometry_table.csv")
```

#Regscore plots - here using the fit objects to contain the fits, not the summaries

#NOTE there is no point using clade as covariate if using pgls, this guarantees a non-significant result because the effect your're looking for is explicitly washed out of the data. But for some reason you feel forced to do pgls or want to play with a different factor, change to procD.pgls function and add "phy=rawtree_clades"

```{r}

#This is a plot of overall allometry without accounting for differences among clades


for (i in 1: length(allometry_clade_gdfs)){
  
  allometry_clade_fits[[i]] <- procD.lm(allometry_clade_gdfs[[i]]$coords ~ log(allometry_clade_gdfs[[i]]$Csize))
      
}

#Making a plot list with the right names
Allom_plots <- allometry_clade_gdfs

for (i in 1: length(allometry_clade_gdfs)){
  
  Allom_plots[[i]]<- plot(allometry_clade_fits[[i]], type="regression", predictor = log(allometry_clade_gdfs[[i]]$Csize))
  
}


```

#Full supplementary Figure plots

```{r}
#Set up colours

skullPlotData_Allom <- skullPlotData

skullPlotData_Allom$PtCol[which(skullPlotData_Allom$Family=="Borhyaenidae")]<-"#e0f0ff"

PlotData_Allom <-skullPlotData_Allom[ match (names(allometry_clade_gdfs$skull$Csize),rownames(skullPlotData_Allom) ),]

pdf(file="./Figures/SuppX_Allometry_partitions.pdf", width=15, height=8)

par(mfrow=c(2,3), mar=c(4,4,3,1), mgp=c(2,0.5,0))

for (i in 1: length(allometry_clade_gdfs)){
  
  plot(Allom_plots[[i]]$RegScore~log(allometry_clade_gdfs[[i]]$Csize),
      xlab= "log Centroid Size of Partition",
      ylab= "Regression Score of Shape ~ log (Csize) Fit",
      main =  names(allometry_clade_fits)[i],
      cex.lab=1.2,
      cex.axis=1.2
         
         )
  
  points(Allom_plots[[i]]$RegScore~log(allometry_clade_gdfs[[i]]$Csize), 
     pch = PlotData_Allom$Point, 
     col = "#000000", 
     bg = PlotData_Allom$PtCol, 
     cex = 2)
  
 

#text(Allom_plots[[i]]$RegScore~log(allometry_clade_gdfs[[i]]$Csize), labels = PlotData_Allom$ID)

  abline(coef(lm(Allom_plots[[i]]$RegScore~log(allometry_clade_gdfs[[i]]$Csize))))

legend("topleft",
       legend = expression("Feliforms", "Caniforms",
                  "Marsupials", "Thylacinidae", italic("C. lupus")),
       pch = c(21, 21, 23, 23),
       col = c("#000000", "#000000", "#000000", "#000000"),
       pt.bg = c("#fff0e0", "#ffaa5c", "#e0f0ff", "#c87ef8", "black"),
       bty = "n",
       pt.cex = 2,
       cex = 1.2,
       text.col = "black",
       horiz = F ,
       inset = c(0.00, 0.0))

}

dev.off()

```



#slope comparison using pairwise
```{r}
#starting with one analysis

#so here I have to re-run the fit because the previous fit objects contained summaries.

#Running pairwise fit for SLOPE differences - relevant for all but neurocranium and MES
allometry_clade_SLOPE_fits <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, Fnp=NULL, Pa=NULL, Mes=NULL)

for (i in 1: length(allometry_clade_gdfs)){
  
  allometry_clade_SLOPE_fits[[i]] <- procD.lm(allometry_clade_gdfs[[i]]$coords ~ log(allometry_clade_gdfs[[i]]$Csize)* allometry_clade_gdfs[[i]]$labels$Clade)
      
}

#Running pairwise fit for SLOPE differences - relevant for all but neurocranium and MES
allometry_clade_MEAN_fits <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, Fnp=NULL, Pa=NULL, Mes=NULL)

for (i in 1: length(allometry_clade_gdfs)){
  
  allometry_clade_MEAN_fits[[i]] <- procD.lm(allometry_clade_gdfs[[i]]$coords ~ log(allometry_clade_gdfs[[i]]$Csize)+ allometry_clade_gdfs[[i]]$labels$Clade)
      
}

#test with just one - works
pairwisecheck <- pairwise(allometry_clade_SLOPE_fits$Cranium, groups=allometry_clade_gdfs[[i]]$labels$Clade, covariate =log(allometry_clade_gdfs$skull$Csize) )
summary (pairwisecheck)

#now the whole thing, going straight for summaries
#these are always the same, so coding them here
groups <- allometry_clade_gdfs$skull$labels$Clade

pairwise_tests_slope <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, FNP=NULL, PA=NULL, MES=NULL)

for (i in 1: length(allometry_clade_SLOPE_fits) ){
  
  pairwise_tests_slope[[i]] <- summary(pairwise(allometry_clade_SLOPE_fits[[i]], groups=groups, covariate =log(allometry_clade_gdfs[[i]]$Csize) ))

}

pairwise_tests_mean <- list(Cranium=NULL, Rostrum=NULL, Neurocranium=NULL, FNP=NULL, PA=NULL, MES=NULL)


for (i in 1: length(allometry_clade_MEAN_fits) ){
  
  pairwise_tests_mean[[i]] <- summary(pairwise(allometry_clade_MEAN_fits[[i]], groups=groups, covariate =log(allometry_clade_gdfs[[i]]$Csize) ))

}

#make table - this is the same for slope and mean results

#Matrix with four rows per comparison (partition name, size, clade and interaction)
pairwise_table_mean <- matrix (nrow=(length(pairwise_tests_mean)*4), ncol = 4)

# Make row names - the below code somehow needs this line to start
rownames(pairwise_table_mean)<- rep (names(pairwise_tests_mean), each=4)

for (i in 2: length(pairwise_tests_mean)){
#need to start off with one  
rownames(pairwise_table_mean)[1:4] <- c(names(pairwise_tests_mean)[[1]], rownames(pairwise_tests_mean[[1]]$summary.table))
#then add to the previous - there must be a nicer way of coding this, but I haven't found it and this is still better than repeated pastes
rownames(pairwise_table_mean)[(((i-1)*4)+1):(i*4)]<- c(names(pairwise_tests_mean)[[i]], rownames(pairwise_tests_mean[[i]]$summary.table))

}

colnames(pairwise_table_mean)<-c("d", "UCL 95%", "Z", "p" )

#Slope table is identical, creating it here
pairwise_table_slope <- pairwise_table_mean

#Now fill in the results, running everything here becasue it's easier and doesn't really matter
#slopes first

for (i in 2: length (pairwise_tests_slope)){
  
#again, 1 needs separate treatment
  #d
  pairwise_table_slope[2:4,1] <- pairwise_tests_slope[[1]]$summary.table$d
  #UCL
  pairwise_table_slope[2:4,2] <- pairwise_tests_slope[[1]]$summary.table$`UCL (95%)`
  #Z
  pairwise_table_slope[2:4,3] <- pairwise_tests_slope[[1]]$summary.table$Z
  #p
  pairwise_table_slope[2:4,4] <- pairwise_tests_slope[[1]]$summary.table$`Pr > d`

  
  #now all the others. i-1*4+2 here (not +1) because the row with the name of the partition is blank
  #SS
  pairwise_table_slope[(((i-1)*4)+2):(i*4),1] <- pairwise_tests_slope[[i]]$summary.table$d
  #Rsq
  pairwise_table_slope[(((i-1)*4)+2):(i*4),2] <- pairwise_tests_slope[[i]]$summary.table$`UCL (95%)`
  #F
  pairwise_table_slope[(((i-1)*4)+2):(i*4),3] <- pairwise_tests_slope[[i]]$summary.table$Z
  #F
  pairwise_table_slope[(((i-1)*4)+2):(i*4),4] <- pairwise_tests_slope[[i]]$summary.table$`Pr > d`
  
}

#now mean
for (i in 2: length (pairwise_tests_mean)){
  
#again, 1 needs separate treatment
  #d
  pairwise_table_mean[2:4,1] <- pairwise_tests_mean[[1]]$summary.table$d
  #UCL
  pairwise_table_mean[2:4,2] <- pairwise_tests_mean[[1]]$summary.table$`UCL (95%)`
  #Z
  pairwise_table_mean[2:4,3] <- pairwise_tests_mean[[1]]$summary.table$Z
  #p
  pairwise_table_mean[2:4,4] <- pairwise_tests_mean[[1]]$summary.table$`Pr > d`

  
  #now all the others. i-1*4+2 here (not +1) because the row with the name of the partition is blank
  #SS
  pairwise_table_mean[(((i-1)*4)+2):(i*4),1] <- pairwise_tests_mean[[i]]$summary.table$d
  #Rsq
  pairwise_table_mean[(((i-1)*4)+2):(i*4),2] <- pairwise_tests_mean[[i]]$summary.table$`UCL (95%)`
  #F
  pairwise_table_mean[(((i-1)*4)+2):(i*4),3] <- pairwise_tests_mean[[i]]$summary.table$Z
  #F
  pairwise_table_mean[(((i-1)*4)+2):(i*4),4] <- pairwise_tests_mean[[i]]$summary.table$`Pr > d`
  
}



write.csv(cbind (pairwise_table_slope,pairwise_table_mean), file="./Figures/Pairwise_tables.csv")

```

