---
title: "01_01_Allometry_check"
author: "Vera Weisbecker"
date: "2024-05-03"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##We are ignoring the effect of allometry on shape because it is not very strong. 
```{r}

library(geomorph)  # CRAN v4.0.7
library(ape)       # CRAN v5.8
library(caper)     # CRAN v1.0.3
library(shapes)    # CRAN v1.2.7
library(tidyverse) # CRAN v2.0.0
library(svglite)   # CRAN v2.1.3
library(here)      # CRAN v1.0.1

load(here("Data", "Processed", "skullLmkSymm.rda"))
load(here("Data", "Processed", "rostralLmkSymm.rda"))
load(here("Data", "Processed", "neuroLmkSymm.rda"))
load(file = here("Data", "Processed", "skullGPA.rda"))


# Load the labels produced in "GMM_GPA.Rmd"
pointLabels <- readRDS(here("Data", "Processed", "pointLabels.rds"))

#Load phylogeny - this is the composite tree from Rovinsky et al., 2021, with the fossil taxa added
# NB: Prothylacinus patagonicus is not included this tree
rawTree <- read.tree(here::here("Data", "phyloGMM.tre"))
```



#Allometry with Andrewsarchus is strong but related to Andrewsarchus being a high-leverage outlier.
```{r}
#shape

new.subset <-coords.subset(A = skullLmkSymm$symm.shape,
                           group = pointLabels$Taxon) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL

#reorder coords to fit the tree
meanCoords <- meanCoords[,,order(match(dimnames(meanCoords)[[3]],rawTree$tip.label))]


Csize <- as.numeric(skullGPA$Csize)
meanSkullCsize <- aggregate(Csize, 
                       by = list(pointLabels$Taxon), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL
rownames(meanSkullCsize) <- meanSkullCsize$Group.1
colnames(meanSkullCsize) <- c("Taxon", "Csize")
meanSkullCsize <- droplevels(meanSkullCsize)

#reorder Csize to fit the tree
meanSkullCsize <- meanSkullCsize[order(match(meanSkullCsize[,1],rawTree$tip.label)),]

#checks
rownames(meanSkullCsize) == rawTree$tip.label
dimnames(meanCoords)[[3]] == rawTree$tip.label
rownames(meanSkullCsize) == dimnames(meanCoords)[[3]]

allometry_gdf <- geomorph.data.frame(coords = meanCoords, Csize = meanSkullCsize$Csize, phy=rawTree)


allometry_fit <- procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdf)

summary(allometry_fit)

#Andrewsarchus is a high-leverage outlier
plot(allometry_fit)


```

#Allometry without Andrewsarchus - residuals are appropriately distributed
```{r}
#take out Andrewsarchus

rawTree <- drop.tip(phy = rawTree, tip = "Andrewsarchus_mongoliensis")

#shape

new.subset <-coords.subset(A = skullLmkSymm$symm.shape[,,-1],
                           group = droplevels( pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL

#reorder coords to fit the tree
meanCoords <- meanCoords[,,order(match(dimnames(meanCoords)[[3]],rawTree$tip.label))]



rawTree$tip.label == dimnames(meanCoords)[[3]]


#Csize without Andrewsarchus
Csize <- as.numeric(skullGPA$Csize)[-1]
meanSkullCsize <- aggregate(Csize, 
                       by = list(pointLabels$Taxon[-1]), 
                       FUN = mean) ##THIS IS NOW ALPHABETICAL

rownames(meanSkullCsize) <- meanSkullCsize$Group.1
colnames(meanSkullCsize) <- c("Taxon", "Csize")
meanSkullCsize <- droplevels(meanSkullCsize)

allometry_gdf <- geomorph.data.frame(coords = meanCoords, Csize = meanSkullCsize$Csize, phy=rawTree)


allometry_fit <- procD.pgls(coords ~ log(Csize), phy=rawTree, data=allometry_gdf)

summary(allometry_fit)

#Here, the residuals are appropriate
plot(allometry_fit)

Allom_PCA <- gm.prcomp(allometry_gdf$coords)


PCA_scores <-  as.data.frame(Allom_PCA$x)
Csize_taxa <- data.frame(Csize=meanSkullCsize$Csize, species = meanSkullCsize$Taxon)
Data_for_pgls <- cbind (PCA_scores, Csize_taxa)

Csize_PCA_frame <- comparative.data( rawTree, Data_for_pgls, names.col = species)

test <- pgls(Comp1~ log(Csize), data=Csize_PCA_frame)


summary(test)

```

#The first two PCAs are not associated with size, then it is smeared across PCs
```{r}
PCs_tested <- 10
pgls_PCA_table <- matrix(nrow=PCs_tested, ncol = 5)
colnames(pgls_PCA_table) <- c("Estimate",  "Std. Error",   "t value", "Rsq", "p")

for (i in 1:PCs_tested){
  
  temp <- summary(pgls( get(paste("Comp", i, sep="")) ~ log(Csize), data=Csize_PCA_frame))
  
 pgls_PCA_table[i,1:3]<- temp$coefficients[2,1:3]
 pgls_PCA_table[i,4]<- temp$r.squared
 pgls_PCA_table[i,5]<- temp$coefficients[2,4]

}

pgls_PCA_table[,c(1:3, 5)] <- round(pgls_PCA_table[,c(1:3, 5)], 3)
pgls_PCA_table[,4] <- round(pgls_PCA_table[,4], 2)

rownames(pgls_PCA_table)<- paste (rep("PC", PCs_tested), c(1:PCs_tested), sep="")

write.csv(pgls_PCA_table, file="../Figures/pgls_PCA_table.csv")


```


#We are justified in keeping Andrewsarchus in the PCA plots because it does not impact on the morphospaces much. The below recreates Fig. 1 without Andrewsarchus.


```{r Labels}
pointLabels$PtCol <- "#fff0e0"
pointLabels$PtCol[pointLabels$Clade == "Metatheria"] <- "#e0f0ff"
pointLabels$PtCol[pointLabels$Family == "Thylacinidae"] <- "#c87ef8" #purple
pointLabels$PtCol[pointLabels$Taxon == "Borhyaena_tuberata"] <- "#ca576c" #red - just to see position. 
pointLabels$PtCol[pointLabels$Family == "Canidae"] <- "#ffaa5c" #blue
pointLabels$PtCol[pointLabels$Taxon == "Canis_lupus"] <- "#233045" #navy - just to see position
pointLabels$PtCol[pointLabels$Taxon == "Andrewsarchus_mongoliensis"] <- "#72c496" #green - just to see position
pointLabels$Point <- 21
pointLabels$Point[pointLabels$Clade == "Metatheria"] <- 23
pointLabels$ID <- str_c(substr(pointLabels$Taxon,1,3), str_to_title(substr(sub(".*_", "", pointLabels$Taxon),1,3)) )
# For publication, we'll change those 'position' colours back to their clade colours and use text call-outs


# Gather mean value labels
meanLabels <- pointLabels
# This pulls out labels by species, and keeps the rest of the associated data columns
meanLabels <- meanLabels %>%
    distinct(Taxon, .keep_all = T)
row.names(meanLabels) <- meanLabels$Taxon
meanLabels <- meanLabels[order(meanLabels$Taxon, decreasing = F), ]
```

#### Skull without Andrewsarchus
```{r Skull Set}

new.subset <-coords.subset(A = skullLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



skullPCA <- gm.prcomp(meanCoords) 
skullPCPlot <- plot(skullPCA)
PCScores <- skullPCA$x[,1:10]
skullPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```

#### Rostrum
```{r Skull Set}
new.subset <-coords.subset(A = rostralLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



rostPCA <- gm.prcomp(meanCoords) 
rostPCPlot <- plot(rostPCA)
PCScores <- rostPCA$x[,1:10]
rostPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```
#### Neurocranium
```{r Skull Set}
new.subset <-coords.subset(A = neuroLmkSymm$symm.shape[,,-1],
                           group = droplevels(pointLabels$Taxon[-1])) ##THIS IS NOW ALPHABETICAL
meanCoords <- sapply(new.subset, 
                      mshape, 
                      simplify = "array") ##THIS IS NOW ALPHABETICAL



neurPCA <- gm.prcomp(meanCoords) 
neurPCPlot <- plot(neurPCA)
PCScores <- neurPCA$x[,1:10]
neurPlotData <- bind_cols(PCScores, meanLabels[-2,])

rm(new.subset, meanCoords, PCScores)

```

#Note that PC2 is flipped but otherwise very similar to Fig. 1
```{r Plot}

pdf(file="../Figures/Supp1_AMong_removed.pdf", width=10, height=10)

par(mfrow=c(2,2))

plot(NULL,
     xlim = range(skullPlotData$Comp1),
     ylim = -range(skullPlotData$Comp2),
     lwd = 0.5,
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Full Cranium")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(skullPlotData$Comp2 ~ skullPlotData$Comp1, 
     pch = skullPlotData$Point, 
     col = "#000000", 
     bg = skullPlotData$PtCol, 
     cex = 2)


#rostrum

plot(NULL,
     xlim = range(rostPlotData$Comp1),
     ylim = -range(rostPlotData$Comp2),
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2 (flipped)",                       
     main = "Rostrum")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(rostPlotData$Comp2 ~ rostPlotData$Comp1, 
     pch = rostPlotData$Point, 
     col = "#000000", 
     bg = rostPlotData$PtCol, 
     cex = 2)



#Neurocranium

plot(NULL,
     xlim = range(neurPlotData$Comp1),
     ylim = range(neurPlotData$Comp2),
     asp = 1,
     xlab = "PC1",                       
     ylab = "PC2",                       
     main = "Neurocranium")
abline(h = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
abline(v = 0, 
       col = "#545454", 
       lty = "dashed", 
       lwd = 0.5)
points(neurPlotData$Comp2 ~ neurPlotData$Comp1, 
     pch = neurPlotData$Point, 
     col = "#000000", 
     bg = neurPlotData$PtCol, 
     cex = 2)

legend("bottomleft",
       legend = expression("Placentals", "Canids",
                  "Metatherians", "Thylacinidae", italic("Borhyaena"), italic("C. lupus")),
       pch = c(21, 21, 23, 23),
       col = c("#000000", "#000000", "#000000", "#000000"),
       pt.bg = c("#fff0e0", "#ffaa5c", "#e0f0ff", "#c87ef8", "#ca576c", "black"),
       bty = "n",
       pt.cex = 2,
       cex = 1,
       text.col = "black",
       horiz = F ,
       inset = c(0.00, 0.0))

plot.new()



dev.off()

```
