---
title: "Rosette_Analysis"
author: "DSRovinsky"
date: "2023-01-17"
output: html_document
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Load packages}
library(geomorph)
library(ape)
library(geiger)
library(nlme)
library(dplyr)
library(svglite)
library(here)
library(Ternary)
library(Rvcg)

# Load phylogeny: this is the composite tree from Rovinsky et al., 2021, with the fossil taxa added
phyTree <- read.tree(here::here("Data", "phyloGMM.tre"))
load(here("Data", "Processed", "rawPoints.rda"))
pointLabels <- readRDS(here("Data", "Processed", "pointLabels.rds"))
```

```{r Functions}
source(here::here("Functions", "Fun.R"))
```


#Creating interlandmark distances: first read in coordinates, then find inter-landmark distances, then average these by species


# matrix of landmarks to measure
# here we are going to measure the distance between lmks 6 & 7, and between lmks 68 & 73
```{r}
lmk_vector <- c(6,7, 68,73,374, 47)
lmks <- matrix(lmk_vector, 
               ncol = 2, 
               byrow = TRUE, 
               dimnames = list(c("Canine", "Pinch", "Height"),
                               c("start", "end")))

# calculate the distances for the  lmks for all specimens
linearDists_allspecies <- interlmkdist(rawPoints$coords, lmks) 

#now average these for each species. I checked a couple of species manually and all is good. Making it readable for the next step but make sure this chunk gets run only once because it's deleting a column!
linearDists <- aggregate(linearDists_allspecies, by=list(pointLabels$Taxon), FUN=mean)
rownames(linearDists) <- linearDists$Group.1
linearDists <- linearDists[,-1]

```

```{r}
# 

# 
#mesh <- vcgPlyRead("./Data/ThCy5749CRA.ply")
# 
# # #a little code to set the cranial orientations for later plotting more easily
# # FOV = 0 # sets parallel projection, rather than orthogonal
# # open3d(); view3d(fov=0);shade3d(mesh) # open a mesh on rgl
# # # # set by hand in rgl window, adjust specimen into dorsal positon then:
# # usrMat_dorsal <- par3d()$userMatrix # save position matrix
# # write.csv(usrMat_dorsal, "./Figures/Usrmats/mesh_usrMat_dorsal.csv",row.names=F) # write to WD for later
# # usrMat_lateral <- par3d()$userMatrix # again, adjust specimen into lateral positon then save
# # write.csv(usrMat_lateral, "./Figures/Usrmats/mesh_usrMat_lateral.csv",row.names=F)
# # usrMat_ventral <- par3d()$userMatrix
# # write.csv(usrMat_ventral, "./Figures/Usrmats/mesh_usrMat_ventral.csv",row.names=F)
# # usrMat_frontal<- par3d()$userMatrix
# # write.csv(usrMat_frontal, "./Figures/Usrmats/mesh_usrMat_frontal.csv",row.names=F)
# 
# 
# # Then you can import for use like:
 usrMat_dorsal <- as.matrix(read.csv("./Figures/Usrmats/mesh_usrMat_dorsal.csv", header = T))
 usrMat_lateral <- as.matrix(read.csv("./Figures/Usrmats/mesh_usrMat_lateral.csv", header = T))
 usrMat_ventral <- as.matrix(read.csv("./Figures/Usrmats/mesh_usrMat_ventral.csv", header = T))
usrMat_frontal <- as.matrix(read.csv("./Figures/Usrmats/mesh_usrMat_frontal.csv", header = T))
# 
# #Just checking that the order and content of point labels and raw points are the same - all good
rawPoints$labels==pointLabels


```

```{r distances figure}

# #mesh and sphere colouration
# #Spheres
#   Colours <- "turquoise"
#   Distballs <- "darkmagenta"
# #Mesh
#   Meshcol <- "mistyrose1"
# #ball radius
#   Radius_base  <- 1.5
# #measurement radius
#   Radius_distpoints <- 2
# #line widths
#   linewd <- 5
# #zoom
#   zoom=0.9
# 
# #pdf(file="./Figures/distances.pdf", width = 8, height = 16)
#   #Overall image specs
# 
# open3d(FOV=0,windowRect=c(0,0,1200,1200), zoom=zoom )
# 
# #set up image with three subscenes in one column
# 
# mfrow3d(2,2)
# par(mar = c(0,0,0,0))
# par(oma = c(0,0,0,0))
# 
# usrmats <- list(usrMat_dorsal, usrMat_lateral,usrMat_ventral,usrMat_frontal)
# 
# 
# for (i in 1:length(usrmats)){
# 
# 
# 
# shade3d (mesh, col= Meshcol, tran = par3d(userMatrix=usrmats[[i]] ), alpha=0.3)#
# 
# spheres3d(rawPoints$coords [,, which(dimnames(rawPoints$coords)[[3]]=="Thylacinus_cynocephalus_5749")], col=Colours, radius=1.5)
# spheres3d(rawPoints$coords [lmk_vector,, which(dimnames(rawPoints$coords)[[3]]=="Thylacinus_cynocephalus_5749")], col=Distballs, radius=2)
# 
# lines3d(rawPoints$coords [lmk_vector[1:2],, which(dimnames(rawPoints$coords)[[3]]=="Thylacinus_cynocephalus_5749")], col=Distballs, lwd=5)
# lines3d(rawPoints$coords [lmk_vector[3:4],, which(dimnames(rawPoints$coords)[[3]]=="Thylacinus_cynocephalus_5749")], col=Distballs, lwd=5)
# lines3d(rawPoints$coords [lmk_vector[5:6],, which(dimnames(rawPoints$coords)[[3]]=="Thylacinus_cynocephalus_5749")], col=Distballs, lwd=5)
# 
# 
#   if (!i == length(usrmats)){next3d()}
#   else {stop}
#   }
# 
# #Now you have to manually zoom in on just the frontal view
# 
# 
# 
# 
# 
# rgl.snapshot(filename = "./Figures/Supp__distances.png")


```
#Here, we're making a dataset of log 10 transformed linear distances for all three - dataRosette
```{r Set up data for analysis}


# get pointLabels for each species
taxaSet <- unique(pointLabels) %>% arrange(Taxon)

#Just checking!
taxaSet$Taxon == rownames(linearDists)

# combine data
dataRosette <- as.data.frame(cbind(taxaSet, log10(linearDists)))
row.names(dataRosette) <- dataRosette$Taxon
# re-order data to tree tip order
dataRosette <- dataRosette[order(match(dataRosette[,3],phyTree$tip.label)),]
name.check(phy = phyTree, 
           data = dataRosette, 
           data.names = dataRosette$Taxon)
plotData <- dataRosette
```

#Here, we create the datasets for prediction interval analysis. the dataPGLS object now contains all data except Andrewsarchus and Thylacinus

```{r PGLS data}
phyPGLS <- drop.tip(phy = phyTree, tip = c("Andrewsarchus_mongoliensis", "Thylacinus_cynocephalus"))
specList <- as.character(phyPGLS$tip.label)
#dataPGLS has no thylacine and Andrewsarchus
dataPGLS <- dataRosette[which(dataRosette$Taxon != c("Andrewsarchus_mongoliensis", "Thylacinus_cynocephalus")),]
dataPGLS <- droplevels(dataPGLS)
name.check(phy = phyPGLS, 
           data = dataPGLS, 
           data.names = dataPGLS$Taxon)

# phylogeny includes fossils (is not ultrametric) so we need to add weights
tW <- diag(vcv(phyPGLS)) # tree weights
```


#These are fits without the thylacine and Andrewsarchus. 
```{r Model fit}
fit1 <- gls(Canine ~ Pinch, 
               data = dataPGLS, 
               correlation = corBrownian(value = 1, 
                                      phy = phyPGLS,
                                      form = ~specList),
               method = "REML",
               weights = varFixed(~tW))

fit2 <- gls(Canine ~ Pinch, 
               data = dataPGLS, 
               correlation = corPagel(value = 1, 
                                      phy = phyPGLS, 
                                      fixed = FALSE,
                                      form = ~specList),
               method = "REML",
               weights = varFixed(~tW))

fit3 <- gls(Canine ~ Pinch, 
               data = dataPGLS, 
               correlation = corBrownian(value = 1, 
                                      phy = phyPGLS,
                                      form = ~specList),
               method = "ML",
               weights = varFixed(~tW))

fit4 <- gls(Canine ~ Pinch, 
               data = dataPGLS, 
               correlation = corPagel(value = 1, 
                                      phy = phyPGLS, 
                                      fixed = FALSE,
                                      form = ~specList),
               method = "ML",
               weights = varFixed(~tW))
anova(fit1, fit2)
anova(fit3, fit4)
```

#These are fits without the thylacine and Andrewsarchus
```{r PGLS Regression}
rosettePGLS_width <- fit4

L = rosettePGLS_width$modelStruct$corStruct[1]

rosetteCI_width <- gls.ci(Y = as.numeric(dataPGLS$Canine),
                    X = as.numeric(dataPGLS$Pinch),
                    Sigma = vcv(corPagel(value = L,
                                       phy = phyPGLS,
                                       fixed = TRUE,
                                       form = ~specList)))


# 95% prediction interval without the thylacine and Andrewsarchus

rosettePI_width <- gls.pi(Y = as.numeric(dataPGLS$Canine),
                    X = as.numeric(dataPGLS$Pinch),
                    Sigma = vcv(corPagel(value = L,
                                       phy = phyPGLS,
                                       fixed = TRUE,
                                       form = ~specList)),
                  k = 2)

```

#These are fits with the whole dataset
```{r Full PGLS}
specList2 <- as.character(phyTree$tip.label)
tWFull <- diag(vcv(phyTree))
fullPGLS_width <- gls(Canine ~ Pinch, 
               data = dataRosette, 
               correlation = corPagel(value = 1, 
                                      phy = phyTree, 
                                      fixed = FALSE,
                                      form = ~specList2),
               method = "ML",
               weights = varFixed(~tWFull))
```



###Same for the height



```{r Model fit}
fit1 <- gls(Height ~ Pinch, 
               data = dataPGLS, 
               correlation = corBrownian(value = 1, 
                                      phy = phyPGLS,
                                      form = ~specList),
               method = "REML",
               weights = varFixed(~tW))

fit2 <- gls(Height ~ Pinch, 
               data = dataPGLS, 
               correlation = corPagel(value = 1, 
                                      phy = phyPGLS, 
                                      fixed = FALSE,
                                      form = ~specList),
               method = "REML",
               weights = varFixed(~tW))

fit3 <- gls(Height ~ Pinch, 
               data = dataPGLS, 
               correlation = corBrownian(value = 1, 
                                      phy = phyPGLS,
                                      form = ~specList),
               method = "ML",
               weights = varFixed(~tW))

fit4 <- gls(Height ~ Pinch, 
               data = dataPGLS, 
               correlation = corPagel(value = 1, 
                                      phy = phyPGLS, 
                                      fixed = FALSE,
                                      form = ~specList),
               method = "ML",
               weights = varFixed(~tW))
anova(fit1, fit2)
anova(fit3, fit4)
```

```{r PGLS Regression}
rosettePGLS_height <- fit4

L = rosettePGLS_height$modelStruct$corStruct[1]

RosetteCI_height <- gls.ci(Y = as.numeric(dataPGLS$Height),
                    X = as.numeric(dataPGLS$Pinch),
                    Sigma = vcv(corPagel(value = L,
                                       phy = phyPGLS,
                                       fixed = TRUE,
                                       form = ~specList)))


# 95% prediction interval without the thylacine and Andrewsarchus

RosettePI_height <- gls.pi(Y = as.numeric(dataPGLS$Height),
                    X = as.numeric(dataPGLS$Pinch),
                    Sigma = vcv(corPagel(value = L,
                                       phy = phyPGLS,
                                       fixed = TRUE,
                                       form = ~specList)),
                  k = 2)

```

```{r Full PGLS}
specList2 <- as.character(phyTree$tip.label)
tWFull <- diag(vcv(phyTree))
fullPGLS_height <- gls(Height ~ Pinch,
               data = dataRosette, 
               correlation = corPagel(value = 1, 
                                      phy = phyTree, 
                                      fixed = FALSE,
                                      form = ~specList2),
               method = "ML",
               weights = varFixed(~tWFull))
```

```{r Plot data}

#The colour scheme is the result of some experimentation and not optimised code-wise for this final published plot
plotData$StemCol <- "#fff0e0" # 20% gray
plotData$StemCol[plotData$Family == "Canidae"] <- "#ffaa5c" # yellow
plotData$StemCol[plotData$Taxon == "Canis_lupus"] <- "#ffaa5c" # navy - to see position
plotData$StemCol[plotData$Clade == "Metatheria"] <- "#e0f0ff" # blue
plotData$StemCol[plotData$Taxon == "Thylacinus_cynocephalus"] <- "#c87ef8" # electric purple
plotData$StemCol[plotData$Taxon == "Nimbacinus_dicksoni"] <- "#e6e6b8" # cream - to see position
plotData$StemCol[plotData$Taxon == "Borhyaena_tuberata"] <- "#fff0e0" # red - to see position
plotData$StemCol[plotData$Taxon == "Andrewsarchus_mongoliensis"] <- "#fff0e0" #  green - to see position

plotData$PtCol <- "#fff0e0" # off-white
plotData$PtCol[plotData$Family == "Canidae"] <- "#ffaa5c" 
plotData$PtCol[plotData$Taxon == "Canis_lupus"] <- "#ffaa5c" # to see position
plotData$PtCol[plotData$Taxon == "Crocuta_crocuta"] <- "#fff0e0" # to see position
plotData$PtCol[plotData$Clade == "Metatheria"] <- "#e0f0ff"
plotData$PtCol[plotData$Taxon == "Thylacinus_cynocephalus"] <- "#c87ef8"
plotData$PtCol[plotData$Taxon == "Nimbacinus_dicksoni"] <- "#c87ef8" # cream - to see position
plotData$PtCol[plotData$Taxon == "Borhyaena_tuberata"] <- "#e0f0ff" # red - to see position
plotData$PtCol[plotData$Taxon == "Andrewsarchus_mongoliensis"] <- "#fff0e0" # green - to see position

plotData$PtSet <- 21
plotData$PtSet[plotData$Clade == "Metatheria"] <- 23

```
#stacked plots!

```{r Regression plot}
 svglite(filename = here::here("Figures", "rosette_Regression_stack.svg"), height = 9, width = 6)

par(mfrow=c(2,1), mar=c(2,4,0,2))



plot(NULL,
    xlim = c(1.2,2), # just a little more so the plot is nice and roomy
     ylim = c(0.8,2.11),
     asp = 1,
     xlab = "Log10 interLmk width at rostral pinch",                       
     ylab = "",  
     #xaxt="n"
     #main = "Rosette PGLS Regression"
     )

title(ylab="Log10 interLmk width at canine", line=2)

polygon(c(rosettePI_width$PI.plot$X, rev (rosettePI_width$PI.plot$X)), 
        c(rosettePI_width$PI.plot$Lower2.5, rev(rosettePI_width$PI.plot$Upper2.5)), 
        col = "#e5e5e5", 
        border = NA )
abline(a = fullPGLS_width$coefficients[1], 
       b = fullPGLS_width$coefficients[2], 
       lty = "longdash",
       lwd = 2, 
       col = "red")
lines(rosetteCI_width$CI.plot$X, 
      rosetteCI_width$CI.plot$Lower2.5,
      lty = 3)
lines(rosetteCI_width$CI.plot$X, 
      rosetteCI_width$CI.plot$Upper2.5,
      lty = 3)
abline(a = rosettePGLS_width$coefficients[1], 
       b = rosettePGLS_width$coefficients[2], 
       lty = "solid",
       lwd = 2, 
       col = "black")
points(y = plotData$Canine, x = plotData$Pinch,
              pch = plotData$PtSet,
              col = "#000000",
              bg = plotData$PtCol,
              cex = 1.5,
              xlab = "Body Mass (Log10 g)",                       
              ylab = "Log10 Centroid size",                       
              main = "Total set Csize")


#This is needed so the polygon doesn't spill out
box(lwd=1)

#############Height. Note you can just re-set the mar() parameters for each individual plot - handy!
#par(mar=c(1,4,0,2))

plot(NULL,
     xlim = c(1.2,2), # just a little more so the plot is nice and roomy
     ylim = c(0.8,2.11), # just a little larger so the plot is nice and roomy
     asp=1,
     xlab = "",                       
     ylab = ""  
     
     )



title(xlab="Log10 interLmk width at rostral 'pinch'", line=2)
title(ylab="Log10 interLmk height at rostral 'pinch'", line=2)



polygon(c(RosettePI_height$PI.plot$X, rev (RosettePI_height$PI.plot$X)), 
        c(RosettePI_height$PI.plot$Lower2.5, rev(RosettePI_height$PI.plot$Upper2.5)), 
        col = "#e5e5e5", 
        border = NA )
abline(a = fullPGLS_height$coefficients[1], 
       b = fullPGLS_height$coefficients[2], 
       lty = "longdash",
       lwd = 2, 
       col = "red")
lines(RosetteCI_height$CI.plot$X, 
      RosetteCI_height$CI.plot$Lower2.5,
      lty = 3)
lines(RosetteCI_height$CI.plot$X, 
      RosetteCI_height$CI.plot$Upper2.5,
      lty = 3)
abline(a = rosettePGLS_height$coefficients[1], 
       b = rosettePGLS_height$coefficients[2], 
       lty = "solid",
       lwd = 2, 
       col = "black")
points(y = plotData$Height, x = plotData$Pinch,
              pch = plotData$PtSet,
              col = "#000000",
              bg = plotData$PtCol,
              cex = 1.5)
#just to find some species
#text(y = plotData$Height, x = plotData$Pinch, labels=plotData$Taxon, cex=0.5)
#text(c(1.7,0.5), "test")

#This is needed so the polygon doesn't spill out
#box(lwd=1)


legend("bottomright",
       legend = c("Placentals", "Canids", "Metatherians",  "Thylacinids"),
       pch = c(21, 21, 23, 23),
       col = "#000000",
       pt.bg = c("#fff0e0", "#ffaa5c", "#e0f0ff", "#c87cf8"),
       bty = "n",
       pt.cex = 1.5,
       cex = 1,
       text.col = "black",
       horiz = F ,
       inset = c(0.01, 0.0), 
       ncol=2, 
       x.intersp=2,
       y.intersp=1.2)
box(col = "#000000")



dev.off()
```


#Trying something new here - ternary plots! Not in ms but fun to do

```{r}


par(mar=c(0,0,0,0))

#OMG this has to be a matrix
Ternary_matrix <- as.matrix(plotData[,c("Canine", "Pinch", "Height")])

TernaryPlot(point = "up", 
            alab = "Canine width", blab = "Pinch width", clab = "Rostrum height",
            region = Ternary_matrix)
                        

#A list, matrix, data.frame or vector in which each element (or row) specifies the three coordinates of a point in ternary space.
TernaryPoints(Ternary_matrix, 
              pch = plotData$PtSet,
              bg = plotData$PtCol,
              cex = 1)
 
#it seems to summarise to me that it there is only one species where height dominates rostral dimensions; in addition, it shares an extremely narrow pinch with Andrewsarchus; as a result, it also has a fairly wide canine width relative to its rostrum height. I do feel that it summarises our conclusion quite well that we've got a uniquely tip-heavy animal here. 

```