---
title: "Weisbecker et al. (2023) marsupial coordinates"
author: "Vera Weisbecker"
date: "26 June 2019"
output: html_document
---

###This is code for the analysis of cranial centroid size in the thylacine compared to other marsupial mammals, using the dataset from "Multiple modes of inference reveal less phylogenetic signal in marsupial basicranial shape compared to the rest of the cranium" by Vera Weisbecker, Leonie Lange-Hodgson, Robin M D Beck, Arianna Harrington, Michael S Y Lee, Thomas Guillerme, and Matthew Phillips, published at the Mammalian Skull Special Issue of the Philosophical Transactions of the Royal Society B. DOI:10.1098/rstb.2022.0085

#It runs a ton of code that was also run in the above paper, but does work to produce Supp. Figure 5.

## load required packages

```{r }

library(geomorph) # CRAN v4.0.10 
library(plyr)     # CRAN v1.8.9      
library(abind)    # CRAN v1.4-8    
library(ape)      # CRAN v5.8-1      
library(here)     # CRAN v1.0.1
library(geiger)   # CRAN v2.0.11
library(nlme)     # CRAN v3.1-168
library(svglite)  # CRAN v2.2.1

# This sets the wd to local
library(rstudioapi) # Safely Access the RStudio API

setwd(dirname(getActiveDocumentContext()$path))


```

## Read in coordinates


```{r }

filelist <- list.files(path= "../Data/Weisbecker_etal_data/Coordinates", pattern = "*.txt")


#Next step is to remove the Museum IDs from the specimen names   
  
 
  names <- gsub (".txt", "", filelist) # extracts names of specimens from the file name
  filelist <- paste("../Data/Weisbecker_etal_data/Coordinates/", filelist, sep="") # rename with path
  coords <- NULL # make empty object that will be filled with 3D array of coordinate data
  for (i in 1:length(filelist)){
    temp  <- read.morphologika(filelist[i]) 
    k <- dim(temp)[1] 
    coords <- rbind(coords, two.d.array(temp)) }
  Data <- arrayspecs(coords, k, 3) 
  dimnames(Data)[[3]] <- names
  remove(i, filelist, k, coords, temp) # clean up environment
  
specimenList <- read.csv("../Data/Weisbecker_etal_data/Species_classifier.csv", header = TRUE, row.names = 1)

length(Data[,,1])

```



```{r}



Landpairs_full <- read.csv("../Data/Weisbecker_etal_data/Landpairs_All.csv", header=FALSE)


```



#GPA for all coordinates

```{r}

#Run GPA
GPA_AllSpecimens <- gpagen(Data, Proj = TRUE, ProcD = FALSE)

#make sure dimnames of coords match with rownames of specdata file
rownames(specimenList) == dimnames(GPA_AllSpecimens$coords)[[3]]



```

#Centroid sizes with _raw_ data
```{r}

#pdf(file="../Figures/SuppFX_centroid_size_checks.pdf",width = 14, height = 14)

par(mfrow=c(2,1))

cexes=0.6

Csize_ordered <- GPA_AllSpecimens$Csize[order(GPA_AllSpecimens$Csize)]

plot( log(Csize_ordered) ~ seq(1:length(Csize_ordered)),pch=19, col="maroon1", xlab= "Specimens ordered by centroid size", ylab = "Centroid size", main = "Specimens ordered by centroid size")
text( log(Csize_ordered) ~ seq(1:length(Csize_ordered)), label = names(Csize_ordered), cex = cexes)

#plot(log(GPA_AllSpecimens$Csize) ~ log (specimenList$Body_mass_g))
#text(log(GPA_AllSpecimens$Csize) ~ log (specimenList$Body_mass_g), labels=rownames(specimenList), cex = cexes)

# svglite(filename = here::here("Figures", "Weisbecker_Csize.svg"),
#         height = 4)
plot(log(GPA_AllSpecimens$Csize) ~  log(specimenList$Body_mass_g), pch=18, col="maroon1", xlab = "Body mass (g)", ylab = "Centroid sizes", main = " Centroid size plotted against body mass")
text(log(GPA_AllSpecimens$Csize) ~  log(specimenList$Body_mass_g), labels=rownames(specimenList), cex = cexes)
abline (lm(log(GPA_AllSpecimens$Csize) ~  log(specimenList$Body_mass_g)))

# dev.off()

remove(cexes, Csize_ordered)

```


#Make a list of species as the basis for the classifier file. I only ran this once, then filled in the remaining data (e.g. diet/loco) manually. This just guarantees that the specimen classifier file and specimens landmarked are definitely the same.

```{r}


#speclist <- matrix(unlist (strsplit(gsub(".txt", "", filelist), "_" )), ncol=3, byrow = TRUE )
#rownames(speclist) <- gsub(".txt", "", filelist); colnames(speclist) <- c("Genus", "Species", "Museum_ID")

#write.csv(speclist, file = "../Data/Raw/Species_classifier.csv")

specimenList$spname <- paste (substr(specimenList$Genus,1,1), substr(specimenList$Species, 1,3), sep = "")

specimenList$Genus_species <- paste (specimenList$Genus, specimenList$Species, sep = "_")

#make sure the specimens and the classifiers are in the same order - this all has to be TRUE
dimnames(GPA_AllSpecimens$coords)[[3]] == rownames (specimenList)

#And average according to species immediately because we don't really need individual values

specieslist <- ddply (specimenList, .(Genus_species), summarize, spname = unique(spname), Genus=unique(Genus), Species = unique(Species), Order=unique(Order), Body_mass_g = mean(Body_mass_g), Locomotion = unique(Locomotion), Diet = unique(Diet))
rownames(specieslist) <- specieslist$Genus_species

#Take away doubled up genus-species designation
specieslist <- specieslist[,-1]

```

#Read in phylogeny (needs to be done here so species orders are OK throughout)

```{r}


raw_tree=read.nexus("../Data/Weisbecker_etal_data/Reference_Phylogeny/Phylogeny_Phillips_merged.nex")

#Changing a few placeholder names on the phylogeny with the species name sampled in reality (for these species, only DNA data from close relatives were available)

raw_tree$tip.label[which(raw_tree$tip.label == "Caluromys_philander")] <- "Caluromys_derbianus"
raw_tree$tip.label[which(raw_tree$tip.label == "Dendrolagus_dorianus")] <- "Dendrolagus_inustus"
raw_tree$tip.label[which(raw_tree$tip.label == "Notoryctes_typhlops")] <- "Notoryctes_caurinus"

namecheck <- geiger::name.check(raw_tree, specieslist)

#Removing all species in the tree that is not apart of the study
tree=drop.tip(raw_tree,namecheck$tree_not_data)

#check that tree and species list are congruent using geiger's name.check
geiger::name.check(tree, specieslist)


#Checking to see if the species coordinates and the species on the tree match and there are no NAs, matchtest length must be 0
Matchtest <- match(tree$tip.label,rownames(specieslist))
length(which(is.na (Matchtest)))

save(tree, file = "../Data/Processed/Tree_processed.rda")

remove(namecheck, raw_tree, Matchtest)

```

#Making sure species list is in the right order for the tree - this can then flow on to the other lists

```{r}

specieslist <- specieslist[match(tree$tip.label,rownames(specieslist)),]

tree$tip.label == rownames(specieslist)


```


#averaging multiple species and removing symmetry

```{r}

# shape data
ind.coords <- aggregate(two.d.array(GPA_AllSpecimens$coords) ~ specimenList$Genus_species, FUN=mean)[,-1]

rownames(ind.coords) <- unique(specimenList$Genus_species)
ind.coords <- arrayspecs(ind.coords, p=ncol(ind.coords)/3, k=3)
#Making sure the coords match up with species lists match up with the phylogeny.
ind.coords <- ind.coords[,,match( rownames(specieslist), dimnames(ind.coords)[[3]])]

dimnames(ind.coords)[[3]] == rownames(specieslist)

#Also just out of paranoia, check a couple of configurations to ensure each species name is allocated to the correct species: 

open3d()
mfrow3d(2,2, sharedMouse=TRUE)

next3d()
plot3d(ind.coords[,,which(dimnames(ind.coords)[[3]]=="Tarsipes_rostratus")], aspect = FALSE,  xlab = "", ylab = "", zlab = "")
decorate3d(main = "Tarsipes - long nose")

next3d()
plot3d(ind.coords[,,which(dimnames(ind.coords)[[3]]=="Vombatus_ursinus")], aspect = FALSE, xlab = "", ylab = "", zlab = "")
decorate3d(main = "Wombat - short nose")

next3d()
plot3d(ind.coords[,,which(dimnames(ind.coords)[[3]]=="Macrotis_lagotis")], aspect = FALSE, xlab = "", ylab = "", zlab = "" )
decorate3d(main = "Bilby - long nose")

next3d()
plot3d(ind.coords[,,which(dimnames(ind.coords)[[3]]=="Petaurus_australis")], aspect = FALSE)
decorate3d( main= "Sugar glider - wide head")

#Now remove asymmetry

GPASymm <- bilat.symmetry(ind.coords, ind=dimnames(ind.coords)[[3]], side=NULL, replicate=NULL, object.sym=TRUE, land.pairs =Landpairs_full)
dimnames(GPASymm$symm.shape)[[3]] <- dimnames(ind.coords)[[3]] # rename

#Are we removing a lot of variation by removing shape? No.
 GPASymm$shape.anova

# Moving on with just the symmetric shape component
ind.coords <- GPASymm$symm.shape 


#Just once, visually making sure the averaging of duplicate specimesn really works - using v. ursinus as an example - looks ok 

#(GPA_AllSpecimens$coords[,,62] + GPA_AllSpecimens$coords[,,61]) /2

### Bilateral symmetry GPA on averaged specimens


# centroid size data, means for multiple specimens
ind.Csize <- as.vector(aggregate(GPA_AllSpecimens$Csize ~ specimenList$Genus_species, FUN=mean)[,-1])
names(ind.Csize) <- unique(specimenList$Genus_species)

ind.Csize <- ind.Csize[match(rownames(specieslist), names(ind.Csize))]



#manually checking just a value for proper averaging, picking V. ursinus and OK 

mean (c(GPA_AllSpecimens$Csize[62], GPA_AllSpecimens$Csize[61])) == ind.Csize[which(names(ind.Csize)=="Vombatus_ursinus")]



#Turning into gdf frame and giving names
GPA_Allmean <- geomorph.data.frame(coords=ind.coords, Csize=ind.Csize)

#We need consensus shapes for later
GPA_allmean_consensus <- GPA_AllSpecimens$consensus


#a few combinations of checks for alignment of species list, phylogeny, csize, and coords. 

dimnames(GPA_Allmean$coords)[[3]] == tree$tip.label;names(GPA_Allmean$Csize) == rownames(specieslist);tree$tip.label == dimnames(GPA_Allmean$coords)[[3]];names(ind.Csize)== tree$tip.label


#Outliers still plot the same - Notoryctes, Vombatus, Tarsipes. Just an additional safety check that the averaging has all worked
plotOutliers(GPA_Allmean$coords)


```

```{r}

```

### Here we change authors and the following poor excuse for code is written by DSRovinsky

```{r}
source(here::here("Functions", "Fun.R"))
```

```{r Organize the data from above}
tree$tip.label == rownames(specieslist)

Csize <- as.data.frame(ind.Csize)
tree$tip.label == rownames(Csize)

specieslist$Csize <- Csize
colnames(specieslist[,8]) <- "Csize"


specList <- as.character(tree$tip.label)

# take the thylacine out, as it is the 'variable' we want to check
specieslistPruned <- specieslist[which(specieslist$Species != "cynocephalus"),]
specieslistPruned <- droplevels(specieslistPruned)

treePruned <- drop.tip(phy = tree, tip = "Thylacinus_cynocephalus")

specListPruned <- specList[which(specList != "Thylacinus_cynocephalus")]
```

```{r Check gls model fits}

# This is the regression without the 'variable' thylacine 
fit1 <- gls(log10(Csize) ~ log10(Body_mass_g), 
            data = specieslistPruned, 
            correlation = corBrownian(value = 1, 
                                      phy = treePruned,
                                      form = ~specListPruned),
            method = "ML")

fit2 <- gls(log10(Csize) ~ log10(Body_mass_g), 
            data = specieslistPruned, 
            correlation = corPagel(value = 1, 
                                      phy = treePruned, 
                                      fixed = FALSE,
                                      form = ~specListPruned),
            method = "ML")

anova(fit1, fit2)
# fit2 model is best
```
```{r qq checks}
plot(fit2, 
     resid(., type = "n")~fitted(.),
     main = "Normalized Residuals v Fitted Values",
     abline = c(0,0), col = "black")
res <- resid(fit2, type="n")
qqnorm(res)
qqline(res)

# looks like Didelphis marsupialis is very much an outlier; we'll leave them in for the time being
```

```{r PGLS CI PI intervals}

L = fit2$modelStruct$corStruct[1]

# 95% confidence interval for the PGLS line

csizeCI <- gls.ci(Y = log10(specieslistPruned$Csize), 
                  X = log10(specieslistPruned$Body_mass_g), 
                  Sigma = vcv(corPagel(value = L, 
                                       phy = treePruned, 
                                       fixed = TRUE, 
                                       form = ~specListPruned)))

# 95% prediction interval for the PGLS line

csizePI <- gls.pi(Y = log10(specieslistPruned$Csize), 
                  X = log10(specieslistPruned$Body_mass_g), 
                  Sigma = vcv(corPagel(value = L, 
                                       phy = treePruned, 
                                       fixed = TRUE,
                                       form = ~specListPruned)),
                  k = 1)
```


```{r PGLS 2}
# Run PGLS with the thylacine included
fullFit1 <- gls(log10(Csize) ~ log10(Body_mass_g), 
            data = specieslist, 
            correlation = corBrownian(value = 1, 
                                      phy = tree,
                                      form = ~specList),
            method = "ML")

fullFit2 <- gls(log10(Csize) ~ log10(Body_mass_g), 
            data = specieslist, 
            correlation = corPagel(value = 1, 
                                      phy = tree, 
                                      fixed = FALSE,
                                      form = ~specList),
            method = "ML")

anova(fullFit1, fullFit2)
# fit2 model is best
```


```{r}
# svglite(filename = here::here("Figures", "MarsupialOnly_Csize_Reg.svg"),
#         height = 5, width = 6)
plot(NULL,
     xlim = range(log10(specieslist$Body_mass_g)),
     ylim = range(log10(specieslist$Csize)), 
     xlab = "Log10 Body Mass (g)",                       
     ylab = "Log10 Csize",  
     main = "PGLS Regression")
polygon(c(csizePI$PI.plot$X, rev (csizePI$PI.plot$X)), 
        c(csizePI$PI.plot$Lower2.5, rev(csizePI$PI.plot$Upper2.5)), 
        col = "#e5e5e5", 
        border = NA )
abline(a = fit2$coefficients[1], 
       b = fit2$coefficients[2], 
       lty = "solid",
       lwd = 2, 
       col = "black")
abline(a = fullFit2$coefficients[1], 
       b = fullFit2$coefficients[2], 
       lty = "longdash",
       lwd = 2, 
       col = "red")
lines(csizeCI$CI.plot$X, 
      csizeCI$CI.plot$Lower2.5,
      lty = 3)
lines(csizeCI$CI.plot$X, 
      csizeCI$CI.plot$Upper2.5,
      lty = 3)
points(y = log10(specieslist$Csize), x = log10(specieslist$Body_mass_g),
              pch = 23,
              col = "#000000",
              bg = "#e0f0ff",
              cex = 1.5,
              xlab = "Log10 Body Mass (g)",                       
              ylab = "Log10 Csize",                       
              main = "PGLS")
lines(x= c(3.5,log10(specieslist$Body_mass_g[which(specieslist$Species =="cynocephalus")])-0.02), 
      y=c(2.6, log10(specieslist$Csize[which(specieslist$Species=="cynocephalus")])-0.02), 
      lwd=1.5 )
text(2.8, 2.6, 
     "Thylacinus cynocephalus", 
     font=3)
lines(x= c(2.5,log10(specieslist$Body_mass_g[which(specieslist$Species =="marsupialis")])-0.02), 
      y=c(2.3, log10(specieslist$Csize[which(specieslist$Species=="marsupialis")])-0.02), 
      lwd=1.5 )
text(1.9, 2.3, 
     "Didelphis marsupialis", 
     font=3)
box(col = "black")
# dev.off()
```

